---
title: "1_DIA_ThreeSpecies_Benchmarking"
author: "Yixin Shi"
date: "2025-08-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(magrittr, ggplot2, ggVennDiagram, pheatmap, ggridges, umap, limma, dplyr, tidyr, viridis, gridExtra, stringr, ggsci, clusterProfiler, org.Mm.eg.db, UpSetR, missForest,ComplexHeatmap, WGCNA, org.Hs.eg.db, GOSemSim, ggrepel, VennDiagram, GSVA, tibble, purrr, scales, imputeLCMD, patchwork, ggpubr)
```

# Working Directory
```{r}
rm(list = ls())
getwd()
```

# Function Settings
```{r}
load_filtered_protein_matrix <- function(file_path) {
  df <- read.csv(file_path, sep = "\t", header = TRUE)
  sample_names <- unlist(lapply(strsplit(colnames(df)[79:94], split = '.', fixed = TRUE), function(x) x[1]))
  colnames(df)[79:94] <- sample_names
  df <- df[,c(3:4,79:94)]
  colnames(df)[1:2] <- c('Protein.Names', 'Genes')
  filter_protein <- df$Protein.Names[!grepl("contam", df$Protein.Names)]
  df <- df[df$Protein.Names %in% filter_protein, ]
  df <- df[, -2]
  rownames(df) <- df$Protein.Names
  df <- df[, -1]
  df <- df[rowSums(df == 0) != ncol(df), ]
  df[df == 0] <- NA
  return(df)
}

extract_metadata_from_colnames <- function(expr_matrix) {
  metadata <- data.frame(
    Sample = colnames(expr_matrix),
    Group  = unlist(lapply(strsplit(colnames(expr_matrix), split = '_', fixed = TRUE), function(x) x[1]))
  )
  rownames(metadata) <- metadata$Sample
  metadata$Group <- factor(metadata$Group)
  levels(metadata$Group) <- c('SampleA', 'SampleB')
  return(metadata)
}

reshape_intensity_long <- function(intensity_matrix, metadata) {
  long_df <- intensity_matrix %>%
    mutate(Protein = rownames(.)) %>%
    pivot_longer(cols = colnames(intensity_matrix),
                 names_to = "Sample",
                 values_to = "Intensity")
  long_df$Group <- plyr::mapvalues(long_df$Sample,
                                   from = metadata$Sample,
                                   to = metadata$Group)
  long_df$Group <- factor(long_df$Group, levels = c(1, 2))
  levels(long_df$Group) <- c("SampleA", "SampleB")
  long_df$Sample <- factor(long_df$Sample, levels = metadata$Sample)
  return(long_df)
}

plot_group_missing_rate <- function(intensity_matrix, long_df, metadata, label = "", nrow_facet = 1) {
  n <- length(unique(rownames(intensity_matrix)))
  aggr <- aggregate(is.na(Intensity) / n ~ Sample, data = long_df, FUN = sum)
  colnames(aggr)[2] <- 'missing_rate'
  aggr$group <- plyr::mapvalues(aggr$Sample,
                                from = metadata$Sample,
                                to = metadata$Group)
  aggr$group <- factor(aggr$group, levels = c(1, 2))
  levels(aggr$group) <- c('SampleA','SampleB')
  global_missing <- sum(is.na(long_df$Intensity))
  global_missing_rate <- global_missing / n / length(metadata$Sample)
  missing_mean <- mean(aggr$missing_rate)
  p <- ggplot(aggr, aes(x = factor(Sample, levels = Sample), y = missing_rate)) +
    geom_bar(stat = 'identity', position = 'dodge', color = "black", fill = "#8AA4C1") +
    geom_hline(yintercept = missing_mean, linetype = 'dashed', color = "red") +
    scale_y_continuous(
      limits = c(0,0.3),
      breaks = seq(0, 0.3, by = 0.05)) +
    xlab("Sample") +
    ylab("Missing rate") +
    ggtitle(paste0("NA ratio (", label, ", out of ", 
                   n, " proteins; Global: ", 
                   round(global_missing_rate, digits = 2), ")")) +
    facet_wrap(~group, nrow = nrow_facet, scales = "free_x") +
    theme_bw() +
    theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_blank(),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    strip.text       = element_text(size = 13, face = "bold") 
  )
  return(p)
}

filter_log2_matrix_by_missing_rate <- function(intensity_matrix, intensity_long, threshold = 0.9) {
  tb_cor <- data.frame(
    row.names = rownames(intensity_matrix),
    missing_rate = rowSums(is.na(intensity_matrix)) / ncol(intensity_matrix),
    abundance = rowMeans(log2(intensity_matrix), na.rm = TRUE)
  )
  intensity_long_log2 <- intensity_long
  intensity_long_log2$Intensity <- log2(intensity_long_log2$Intensity)
  valid_proteins <- rownames(tb_cor)[tb_cor$missing_rate < threshold]
  long_filtered <- intensity_long_log2 %>%
    filter(Protein %in% valid_proteins)
  wide_filtered <- long_filtered %>%
    dplyr::select(-Group) %>%
    pivot_wider(names_from = Sample, values_from = Intensity) %>%
    as.data.frame()
  rownames(wide_filtered) <- wide_filtered$Protein
  wide_filtered <- wide_filtered[, -1]  
  return(wide_filtered)
}

median_centering <- function(df) {
  df_norm <- as.data.frame(apply(df, 2, function(x) {
    gMedian = median(x, na.rm = TRUE)
    x - gMedian
  }
  ))
  return(df_norm)
}

run_minprob <- function(data_list, 
                              q = 0.01, 
                              tune.sigma = 1, 
                              seed = 25, 
                              save_dir = "./output_DDA", 
                              prefix = "minProb") {
  if (!dir.exists(save_dir)) dir.create(save_dir, recursive = TRUE)
  results <- list()
  for (name in names(data_list)) {
    cat("Running minProb imputation for:", name, "\n")
    set.seed(seed)
    data <- data_list[[name]]
    time_taken <- system.time({
      result <- imputeLCMD::impute.MinProb(data, q = q, tune.sigma = tune.sigma)
    })
    save_path <- file.path(save_dir, paste0(prefix, "_", name, ".Rdata"))
    save(result, file = save_path)
    results[[name]] <- list(
      imputed_data = result,
      time = time_taken,
      save_path = save_path
    )
    print(time_taken)
  }
  return(results)
}

run_gaussian_imputation <- function(df, 
                width = 0.3, 
                downshift = 1.8, 
                seed = 25, 
                save_path = NULL) {
  set.seed(seed)
  observed_values <- unlist(df)[!is.na(unlist(df))]
  mu <- mean(observed_values)
  sigma <- sd(observed_values)
  impute_mean <- mu - downshift * sigma
  impute_sd <- sigma * width
  time_taken <- system.time({
    imputed_df <- df
    na_indices <- which(is.na(df), arr.ind = TRUE)
    imputed_df[na_indices] <- rnorm(n = nrow(na_indices), mean = impute_mean, sd = impute_sd)
  })
  if (!is.null(save_path)) {
    save(imputed_df, file = save_path)
  }
  print(time_taken)
  return(list(
    imputed_data = imputed_df,
    time = time_taken,
    save_path = save_path
  ))
}

run_gaussian <- function(data_list, 
                               width = 0.3, 
                               downshift = 1.8, 
                               seed = 25, 
                               save_dir = "./output_DDA", 
                               prefix = "GS") {
  if (!dir.exists(save_dir)) dir.create(save_dir, recursive = TRUE)
  results <- list()
  for (name in names(data_list)) {
    cat("Running Gaussian imputation for:", name, "\n")
    save_path <- file.path(save_dir, paste0(prefix, "_", name, ".Rdata"))
    result <- run_gaussian_imputation(
      df = data_list[[name]],
      width = width,
      downshift = downshift,
      seed = seed,
      save_path = save_path
    )
    results[[name]] <- result
  }
  return(results)
}

run_knn_imputation <- function(df, 
                               k = 10, 
                               seed = 25, 
                               save_path = NULL) {
  if (!requireNamespace("impute", quietly = TRUE)) {
    stop("Please install the 'impute' package.")
  }
  set.seed(seed)
  time_taken <- system.time({
    knn_result <- impute::impute.knn(as.matrix(df), k = k)
    knn_df <- as.data.frame(knn_result$data)
  })
  if (!is.null(save_path)) {
    save(knn_df, file = save_path)
  }
  print(time_taken)
  return(list(
    imputed_data = knn_df,
    time = time_taken,
    save_path = save_path
  ))
}

run_knn <- function(data_list, 
                    k = 10, 
                    seed = 25, 
                    save_dir = "./output_DDA", 
                    prefix = "knn") {
  if (!dir.exists(save_dir)) dir.create(save_dir, recursive = TRUE)
  results <- list()
  for (name in names(data_list)) {
    cat("Running KNN imputation for:", name, "\n")
    save_path <- file.path(save_dir, paste0(prefix, "_", name, ".Rdata"))
    result <- run_knn_imputation(
      df = data_list[[name]],
      k = k,
      seed = seed,
      save_path = save_path
    )
    results[[name]] <- result
  }
  return(results)
}

run_rf_imputation <- function(df, 
                              seed = 123, 
                              save_path = NULL, 
                              verbose = FALSE) {
  if (!requireNamespace("missForest", quietly = TRUE)) {
    stop("Please install the 'missForest' package.")
  }
  set.seed(seed)
  time_taken <- system.time({
    res <- missForest::missForest(as.data.frame(df), verbose = verbose)
    imputed_df <- res$ximp
  })
  if (!is.null(save_path)) {
    save(imputed_df, file = save_path)
  }
  print(time_taken)
  return(list(
    imputed_data = imputed_df,
    time = time_taken,
    save_path = save_path
  ))
}

run_rf <- function(data_list, 
                   seed = 123, 
                   save_dir = "./output_DDA", 
                   prefix = "rf", 
                   verbose = FALSE) {
  if (!dir.exists(save_dir)) dir.create(save_dir, recursive = TRUE)
  results <- list()
  for (name in names(data_list)) {
    cat("Running missForest imputation for:", name, "\n")
    save_path <- file.path(save_dir, paste0(prefix, "_", name, ".Rdata"))
    result <- run_rf_imputation(
      df = data_list[[name]],
      seed = seed,
      save_path = save_path,
      verbose = verbose
    )
    results[[name]] <- result
  }
  return(results)
}

```
## Soft Hybrid Imputation Function
```{r}
soft_hybrid_impute <- function(df_rf, df_minProb, df_raw, 
                               r0 = NULL, x0 = NULL, a = 10, b = 5, lambda = 0.4,
                               visualize = TRUE, group_name = NULL,
                               save_dir = "./SoftHybrid_DIA") {
  stopifnot(all(dim(df_rf) == dim(df_minProb)))
  stopifnot(all(rownames(df_rf) == rownames(df_minProb)))

  missing_rate <- rowMeans(is.na(df_raw))
  mean_intensity <- rowMeans(df_raw, na.rm = TRUE)
  miss_cor <- data.frame(missing_rate = missing_rate, mean_abundance = mean_intensity)

  q3 <- quantile(miss_cor$missing_rate, 0.75, na.rm = TRUE)
  skip_loess <- q3 == 0

  if (skip_loess) {
    elbow_x <- 1 / ncol(df_raw)
    candidates <- mean_intensity[round(missing_rate, 6) == round(elbow_x, 6)]
    if (length(candidates) == 0) {
      elbow_y <- mean(mean_intensity, na.rm = TRUE)  
    } else {
      elbow_y <- mean(candidates, na.rm = TRUE)
    }
  } else {
    loess_fit <- loess(mean_abundance ~ missing_rate, data = miss_cor, span = 0.75)
    x_vals <- seq(min(miss_cor$missing_rate), max(miss_cor$missing_rate), length.out = 1000)
    y_vals <- predict(loess_fit, newdata = data.frame(missing_rate = x_vals))
    x_start <- x_vals[1]; y_start <- y_vals[1]
    x_end   <- x_vals[length(x_vals)]; y_end <- y_vals[length(y_vals)]
    line_vec <- c(x_end - x_start, y_end - y_start)
    distances <- sapply(1:length(x_vals), function(i) {
      point_vec <- c(x_vals[i] - x_start, y_vals[i] - y_start)
      cross_prod <- abs(line_vec[1] * point_vec[2] - line_vec[2] * point_vec[1])
      norm_line <- sqrt(sum(line_vec^2))
      return(cross_prod / norm_line)
    })
    max_index <- which.max(distances)
    elbow_x <- x_vals[max_index]
    elbow_y <- y_vals[max_index]
  }
  r0 <- elbow_x
  x0 <- elbow_y
  sigmoid <- function(x, k, x0) {
    1 / (1 + exp(-k * (x - x0)))
  }
  sigmoid_r <- sigmoid(missing_rate, k = a, x0 = r0)
  sigmoid_x <- sigmoid(mean_intensity, k = b, x0 = x0)
  p_MNAR <- sigmoid_r * (1 - lambda * sigmoid_x)
  df_imp_soft <- df_rf
  for (i in seq_len(nrow(df_rf))) {
    w <- 1 - p_MNAR[i]
    df_imp_soft[i, ] <- w * df_rf[i, ] + (1 - w) * df_minProb[i, ]
  }
  if (visualize) {
    library(ggplot2)
    df_plot <- data.frame(
      Protein = rownames(df_rf),
      MissingRate = missing_rate,
      MeanIntensity = mean_intensity,
      Weight_RF = 1 - p_MNAR
    )
    p <- ggplot(df_plot, aes(x = MissingRate, y = MeanIntensity, color = Weight_RF)) +
    geom_point(size = 1.6, alpha = 0.8) +
    scale_color_gradient(low = "blue", high = "red", limits = c(0, 1)) +
    scale_x_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, by = 0.25),
      labels = percent_format(accuracy = 1)
    ) +
    scale_y_continuous(expand = expansion(mult = c(0.02, 0.05))) +
    labs(
      title = paste0("Soft Hybrid Weights (", group_name, ")"),
      x = "Missing Rate",
      y = "Mean Intensity",
      color = "RF Weight"
    ) +
    theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold", angle = 45, hjust = 1),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  ) +
    guides(color = guide_colorbar(
      ticks = TRUE,
      barheight = unit(48, "pt"),
      frame.colour = "black"
    ))
    print(p)
    if (!dir.exists(save_dir)) dir.create(save_dir)
    save(p, file = paste0(save_dir, '/', group_name,"_SoftHybrid_plot.Rdata"))
    ggsave(
      filename = file.path(save_dir, paste0(group_name,"_SoftHybrid_plot.tiff")), 
         plot = p, 
         device = "tiff",
         dpi = 300,
         width = 6, 
         height = 3.5,
         units = "in"
    )
  }
  return(df_imp_soft)
}

run_soft_hybrid_all_groups <- function(raw_list, rf_list, minProb_list,
                                       lambda = 0.4, a = 10, b = 5, visualize = TRUE) {
  group_names <- names(raw_list)
  result_list <- list()
  for (group in group_names) {
    cat("Processing group:", group, "\n")
    df_raw <- raw_list[[group]]
    df_rf <- rf_list[[group]][['imputed_data']]
    df_minProb <- minProb_list[[group]][['imputed_data']]
    result <- soft_hybrid_impute(df_rf, df_minProb, df_raw,
                                 a = a, b = b, lambda = lambda, visualize = visualize,
                                 group_name = group)
    result_list[[group]] <- result
  }
  return(result_list)
}
```
# Data Preperation
## Read Data
```{r, warning=FALSE}
intensity_50pg <- load_filtered_protein_matrix("./data/combined_protein_50pg.tsv")
intensity_100pg <- load_filtered_protein_matrix("./data/combined_protein_100pg.tsv")
intensity_1ng <- load_filtered_protein_matrix("./data/combined_protein_1ng.tsv")
intensity_10ng <- load_filtered_protein_matrix("./data/combined_protein_10ng.tsv")
```
## Metadata Preparation
```{r}
metadata_50pg  <- extract_metadata_from_colnames(intensity_50pg)
metadata_100pg <- extract_metadata_from_colnames(intensity_100pg)
metadata_1ng   <- extract_metadata_from_colnames(intensity_1ng)
metadata_10ng  <- extract_metadata_from_colnames(intensity_10ng)
```
## Wide to Long
```{r}
intensity_50pg_long <- reshape_intensity_long(intensity_50pg, metadata_50pg)
intensity_100pg_long <- reshape_intensity_long(intensity_100pg, metadata_100pg)
intensity_1ng_long <- reshape_intensity_long(intensity_1ng, metadata_1ng)
intensity_10ng_long <- reshape_intensity_long(intensity_10ng, metadata_10ng)
```
## Missing Rate Calculation and Filtration
```{r}
df_50pg_filter <- filter_log2_matrix_by_missing_rate(intensity_50pg, intensity_50pg_long)
df_100pg_filter <- filter_log2_matrix_by_missing_rate(intensity_100pg, intensity_100pg_long)
df_1ng_filter <- filter_log2_matrix_by_missing_rate(intensity_1ng, intensity_1ng_long)
df_10ng_filter <- filter_log2_matrix_by_missing_rate(intensity_10ng, intensity_10ng_long)
```
## Missing Rate Presentation
```{r}
p1 <- plot_group_missing_rate(intensity_50pg, intensity_50pg_long, metadata_50pg, label = "50pg_DDA")
p2 <- plot_group_missing_rate(intensity_100pg, intensity_100pg_long, metadata_100pg, label = "100pg_DDA")
p3 <- plot_group_missing_rate(intensity_1ng, intensity_1ng_long, metadata_1ng, label = "1ng_DDA")
p4 <- plot_group_missing_rate(intensity_10ng, intensity_10ng_long, metadata_10ng, label = "10ng_DDA")

p1
p2
p3
p4

pcombo <- (p1 | p2) / (p3 | p4) +   
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') &   
  theme(
    plot.tag = element_text(face = "bold", size = 14),
    plot.tag.position = c(0.02, 0.98)         
  )

pcombo

ggsave(
  filename = "./figure/DDA_MissingRate_ThreeSpecies.tiff",
  plot = pcombo,                      
  device = "tiff",                    
  dpi = 300,                          
  width = 15,                        
  height = 6,                       
  units = "in"                      
)
```
## Normalisation
```{r}
df_50pg_norm <- median_centering(df_50pg_filter)
df_100pg_norm <- median_centering(df_100pg_filter)
df_1ng_norm <- median_centering(df_1ng_filter)
df_10ng_norm <- median_centering(df_10ng_filter)
```

# Data Imputation
```{r}
data_list <- list(
  "50pg"  = df_50pg_norm,
  "100pg" = df_100pg_norm,
  "1ng"   = df_1ng_norm,
  "10ng"  = df_10ng_norm
)
data_filter_list <- list(
  "50pg"  = df_50pg_filter,
  "100pg" = df_100pg_filter,
  "1ng"   = df_1ng_filter,
  "10ng"  = df_10ng_filter
)
```

## Elbow Point Detection
```{r}
theme_pub <- function(base_size = 10, base_family = "Arial") {
  ggplot2::theme_classic(base_size = base_size, base_family = base_family) %+replace%
    ggplot2::theme(
      panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
      axis.title       = element_text(size = base_size + 1, face = "bold"),
      axis.text        = element_text(size = base_size),
      axis.ticks.length= unit(2, "mm"),
      plot.title       = element_text(size = base_size + 2, face = "bold", hjust = 0.5),
      legend.position  = "right",
      legend.title     = element_text(size = base_size),
      legend.text      = element_text(size = base_size - 1)
    )
}

find_missing_rate_elbow <- function(df_norm,
                                    span = 0.75,
                                    n_points = 1000,
                                    point_size = 1.6,
                                    point_alpha = 0.1,
                                    save_path = NULL, 
                                    group_name = NULL,
                                    tiff_width = 8,  
                                    tiff_height = 5,
                                    tiff_dpi = 600) {
  
  stopifnot(is.matrix(df_norm) || is.data.frame(df_norm))
  if (!requireNamespace("ggplot2", quietly = TRUE)) stop("Please install 'ggplot2'.")
  if (!requireNamespace("scales", quietly = TRUE))  stop("Please install 'scales'.")
  
  df_norm <- as.matrix(df_norm)
  
  miss_cor <- data.frame(
    missing_rate   = rowSums(is.na(df_norm)) / ncol(df_norm),
    mean_abundance = rowMeans(df_norm, na.rm = TRUE)
  )
  
  fit <- stats::loess(mean_abundance ~ missing_rate, data = miss_cor, span = span)
  
  x_vals <- seq(min(miss_cor$missing_rate, na.rm = TRUE),
                max(miss_cor$missing_rate, na.rm = TRUE),
                length.out = n_points)
  y_vals <- stats::predict(fit, newdata = data.frame(missing_rate = x_vals))
  
  keep   <- !is.na(y_vals)
  x_vals <- x_vals[keep]
  y_vals <- y_vals[keep]
  
  x_start <- x_vals[1];              y_start <- y_vals[1]
  x_end   <- x_vals[length(x_vals)]; y_end   <- y_vals[length(y_vals)]
  line_vec  <- c(x_end - x_start, y_end - y_start)
  norm_line <- sqrt(sum(line_vec^2))
  distances <- vapply(seq_along(x_vals), function(i) {
    point_vec <- c(x_vals[i] - x_start, y_vals[i] - y_start)
    abs(line_vec[1] * point_vec[2] - line_vec[2] * point_vec[1]) / norm_line
  }, numeric(1))
  max_index <- which.max(distances)
  elbow_x <- x_vals[max_index]
  elbow_y <- y_vals[max_index]
  
  library(ggplot2)
  library(scales)
  
  p <- ggplot(miss_cor, aes(x = missing_rate, y = mean_abundance)) +
    geom_point(size = point_size, alpha = point_alpha) +
    geom_smooth(method = "loess", se = FALSE, span = span) +
    geom_abline(
      slope = (y_end - y_start) / (x_end - x_start),
      intercept = y_start - ((y_end - y_start) / (x_end - x_start)) * x_start,
      linetype = "dashed"
    ) +
    annotate("point", x = elbow_x, y = elbow_y, colour = "red", size = 2) +
    annotate("text",
             x = elbow_x, y = elbow_y + 2,
             label = paste0("Elbow @ ", scales::percent(elbow_x, accuracy = 0.01)),
             hjust = -0.2, vjust = -0.5, colour = "red",
             size = 4.5, fontface = "bold") +
    scale_x_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, by = 0.25),
      labels = scales::percent_format(accuracy = 1)
    ) +
    scale_y_continuous(
      limits = c(-5,10),
      breaks = seq(-5, 10, by = 2.5),
      expand = expansion(mult = c(0.02, 0.05))) +
    labs(
      title = paste("Loess-Based Elbow Point by Maximum Distance Method:", group_name),
      x = "Missing Rate",
      y = "Mean log2 Abundance"
    ) +
    theme_bw() +
    theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
  
  if (!is.null(save_path)) {
    ggsave(filename = save_path, plot = p, device = "tiff",
           dpi = tiff_dpi, width = tiff_width, height = tiff_height,
           units = "in", compression = "lzw")
  }
  
  list(
    miss_cor = miss_cor,
    fit = fit,
    elbow = list(x = elbow_x, y = elbow_y),
    plot = p
  )
}
```

```{r}
res_50pg_DDA <- find_missing_rate_elbow(df_50pg_norm, span = 0.75,
                                        group_name = 'DDA_50pg',
                                        save_path = "./figure/DDA_50pg_elbow.tiff")
res_100pg_DDA <- find_missing_rate_elbow(df_100pg_norm, span = 0.75,
                                         group_name = 'DDA_100pg',
                                         save_path = "./figure/DDA_100pg_elbow.tiff")
res_1ng_DDA <- find_missing_rate_elbow(df_1ng_norm, span = 0.75,
                                       group_name = 'DDA_1ng',
                                       save_path = "./figure/DDA_1ng_elbow.tiff")
res_10ng_DDA <- find_missing_rate_elbow(df_10ng_norm, span = 0.75,
                                        group_name = 'DDA_10ng',
                                        save_path = "./figure/DDA_10ng_elbow.tiff")
save(res_50pg_DDA, file = './output/res_50pg_DDA.Rdata')
save(res_100pg_DDA, file = './output/res_10pg_DDA.Rdata')
save(res_1ng_DDA, file = './output/res_1ng_DDA.Rdata')
save(res_10ng_DDA, file = './output/res_10ng_DDA.Rdata')
```

```{r}
load(file = './output/res_50pg_DDA.Rdata')
load(file = './output/res_10pg_DDA.Rdata')
load(file = './output/res_1ng_DDA.Rdata')
load(file = './output/res_50pg_DIA.Rdata')
load(file = './output/res_100pg_DIA.Rdata')
load(file = './output/res_1ng_DIA.Rdata')
```

```{r, warning=FALSE}
pelbow_combo <-
 (res_50pg_DIA$plot|res_100pg_DIA$plot|res_1ng_DIA$plot)  / (res_50pg_DDA$plot|res_100pg_DDA$plot|res_1ng_DDA$plot) +   
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') & 
  theme(
    plot.tag = element_text(face = "bold", size = 14),
    plot.tag.position = c(0.02, 0.98)         
  )

ggsave(
  filename = "./figure/Elbow_Combo.tiff",  
  plot = pelbow_combo,                      
  device =  'tiff',                    
  dpi = 300,                          
  width = 22,                         
  height = 11,                        
  units = "in"                        
)
pelbow_combo
```


## Imputation Method I: Minimal Probability of ALL proteins
```{r}
#minprob_results <- run_minprob(data_list)
#save(minprob_results, file = './output_DDA/minprob_results.Rdata')
load(file = './output_DDA/minprob_results.Rdata')
```
## Imputation Method II: Gaussian sample of ALL proteins
```{r}
#gaussian_results <- run_gaussian(data_list)
#save(gaussian_results, file = './output_DDA/gaussian_results.Rdata')
load(file = './output_DDA/gaussian_results.Rdata')
```
## Imputation Method III: KNN of ALL proteins
```{r}
#knn_results <- run_knn(data_list)
#save(knn_results, file = './output_DDA/knn_results.Rdata')
load(file = './output_DDA/knn_results.Rdata')
```
## Imputation Method IV: Random Forest of ALL proteins
```{r}
#rf_results <- run_rf(data_list)
#save(rf_results, file = './output_DDA/rf_results.Rdata')
load(file = './output_DDA/rf_results.Rdata')
```
## Imputation Method V: Soft Hybrid Imputation
```{r}
imp_soft_list <- run_soft_hybrid_all_groups(data_list, rf_results, minprob_results,lambda = 0.4, a = 10, b = 5, visualize = TRUE)
softhybrid_results <- list()
softhybrid_results[['50pg']]$imputed_data <- imp_soft_list[['50pg']]
softhybrid_results[['100pg']]$imputed_data <- imp_soft_list[['100pg']]
softhybrid_results[['1ng']]$imputed_data <- imp_soft_list[['1ng']]
softhybrid_results[['10ng']]$imputed_data <- imp_soft_list[['10ng']]
save(softhybrid_results, file = './output_DDA/softhybrid_results.Rdata')
load(file = './output_DDA/softhybrid_results.Rdata')
```

```{r}
load(file = './SoftHybrid_DIA/10ng_SoftHybrid_plot.Rdata')
pDIA_10ng_SH <- p + theme_bw()
load(file = './SoftHybrid_DIA/1ng_SoftHybrid_plot.Rdata')
pDIA_1ng_SH <- p + theme_bw()
load(file = './SoftHybrid_DIA/100pg_SoftHybrid_plot.Rdata')
pDIA_100pg_SH <- p + theme_bw()
load(file = './SoftHybrid_DIA/50pg_SoftHybrid_plot.Rdata')
pDIA_50pg_SH <- p + theme_bw()
load(file = './SoftHybrid_DDA/10ng_SoftHybrid_plot.Rdata')
pDDA_10ng_SH <- p + theme_bw()
load(file = './SoftHybrid_DDA/1ng_SoftHybrid_plot.Rdata')
pDDA_1ng_SH <- p + theme_bw()
load(file = './SoftHybrid_DDA/100pg_SoftHybrid_plot.Rdata')
pDDA_100pg_SH <- p + theme_bw()
load(file = './SoftHybrid_DDA/50pg_SoftHybrid_plot.Rdata')
pDDA_50pg_SH <- p + theme_bw()
``` 

```{r, warning=FALSE}
pSH_DIA <- (pDIA_50pg_SH|pDIA_100pg_SH|pDIA_1ng_SH|pDIA_10ng_SH) +   
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') &   
  theme(
    plot.tag = element_text(face = "bold", size = 14),
    plot.tag.position = c(0.02, 0.98)         
  ) +
    theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pSH_DIA

pSH_DDA <- (pDDA_50pg_SH|pDDA_100pg_SH|pDDA_1ng_SH|pDDA_10ng_SH) +   
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') &   
  theme(
    plot.tag = element_text(face = "bold", size = 14),
    plot.tag.position = c(0.02, 0.98)         
  ) +
    theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pSH_DDA
```

```{r}
pSH_combo <-
  (pSH_DIA /pSH_DDA)+   
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') &   
  theme(
    plot.tag = element_text(face = "bold", size = 14),
    plot.tag.position = c(0.02, 0.98)         
  ) +
    theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )

ggsave(
  filename = "./figure/SH_Combo.tiff",  
  plot = pSH_combo,                     
  device = "tiff",                    
  dpi = 300,                          
  width = 18,                         
  height = 9,                        
  units = "in"                        
)

pSH_combo
```

# Accuracy Benchmarking
```{r}
plot_log2_ratio <- function(expr_mat, metadata, raw_data, title_prefix, save_dir = "./plots_DDA") {
  groupA <- metadata$Sample[metadata$Group == "SampleA"]
  groupB <- metadata$Sample[metadata$Group == "SampleB"]
  mean_A <- rowMeans(expr_mat[, as.character(groupA)])
  mean_B <- rowMeans(expr_mat[, as.character(groupB)])
  intensity_B <- rowMeans(raw_data[, as.character(groupB)])
  log2_ratio <- mean_A - mean_B
  species <- rep(NA, nrow(expr_mat))
  species[grepl("_HUMAN$", rownames(expr_mat))] <- "Human"
  species[grepl("_YEAST$", rownames(expr_mat))] <- "Yeast"
  species[grepl("_ECOLI$|_ECOBD$|_ECOL6$", rownames(expr_mat))] <- "Ecoli"
  species[is.na(species)] <- "Other"
  plot_df <- data.frame(
    Protein = rownames(expr_mat),
    Mean_B = intensity_B,
    Log2_Ratio = log2_ratio,
    Species = species
  )
  if (!dir.exists(save_dir)) dir.create(save_dir)
  custom_colors <- c("Human" = "#1b9e77", "Yeast" = "#d95f02", "Ecoli" = "#7570b3", "Other" = "grey70")
  p <- ggplot(plot_df, aes(x = Mean_B, y = Log2_Ratio, color = Species)) +
    geom_point(size = 2, alpha = 0.6) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.5) +
    geom_hline(yintercept = c(-2, 1), linetype = "dashed", color = "black", linewidth = 0.5) +
    coord_cartesian(ylim = c(-3, 3), expand = FALSE) +
    scale_color_manual(values = custom_colors) +
    labs(
      title = title_prefix,
      x = "Log2(SampleB)",
      y = "Log2(SampleA/SampleB)"
    ) +
    theme_bw() +
    theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
  ggsave(filename = file.path(save_dir, paste0(title_prefix, "_plot.tiff")), 
         plot = p, 
         device = "tiff",
         dpi = 300,
         width = 6, 
         height = 5,
         units = "in")
  save(p, file = paste0(save_dir, '/', title_prefix, '.Rdata'))
  return(plot_df)
}
```

```{r}
impute_methods <- c("minProb", "RF", "Gaussian", "KNN", 'SoftHybrid')
doses <- c("50pg", "100pg", "1ng", "10ng")
results_list <- list()
results_list_raw <- list()
```

```{r}
for (method in impute_methods) {
  for (dose in doses) {
    expr_mat <- get(paste0(tolower(method), "_results"))[[dose]]$imputed_data
    raw_data <- get(paste0("df_", dose, "_filter"))
    metadata <- get(paste0("metadata_", dose))
    title_prefix <- paste0(method, "-", dose)
    
    res <- plot_log2_ratio(expr_mat, metadata, raw_data, title_prefix)
    results_list[[title_prefix]] <- res
  }
}
for (dose in doses) {
    expr_mat <- get(paste0("df_", dose, "_filter"))
    raw_data <- get(paste0("df_", dose, "_filter"))
    metadata <- get(paste0("metadata_", dose))
    title_prefix <- paste0("Raw_", dose)
    
    res <- plot_log2_ratio(expr_mat, metadata, raw_data, title_prefix)
    results_list_raw[[title_prefix]] <- res
}
```

```{r}
save(results_list, file = './data/results_list_DDA.Rdata')
save(results_list_raw, file = './data/results_list_raw_DDA.Rdata')
```

```{r}
results_list_DDA <- results_list
results_list_raw_DDA <- results_list_raw

load(file = './data/results_list_DIA.Rdata')
load(file = './data/results_list_raw_DIA.Rdata')
```
## ID number bar plot
```{r}
ID_50pg_raw_DIA <- sum(!is.na(results_list_raw$Raw_50pg$Log2_Ratio))
ID_50pg_raw_DDA <- sum(!is.na(results_list_raw_DDA$Raw_50pg$Log2_Ratio))
ID_100pg_raw_DIA <- sum(!is.na(results_list_raw$Raw_100pg$Log2_Ratio))
ID_100pg_raw_DDA <- sum(!is.na(results_list_raw_DDA$Raw_100pg$Log2_Ratio))
ID_1ng_raw_DIA <- sum(!is.na(results_list_raw$Raw_1ng$Log2_Ratio))
ID_1ng_raw_DDA <- sum(!is.na(results_list_raw_DDA$Raw_1ng$Log2_Ratio))
ID_10ng_raw_DIA <- sum(!is.na(results_list_raw$Raw_10ng$Log2_Ratio))
ID_10ng_raw_DDA <- sum(!is.na(results_list_raw_DDA$Raw_10ng$Log2_Ratio))

df_bar <- data.frame(
  'Method' = c('DIA_raw_50pg', 'DIA_imputed_50pg', 'DDA_raw_50pg', 'DDA_imputed_50pg',
               'DIA_raw_100pg', 'DIA_imputed_100pg', 'DDA_raw_100pg', 'DDA_imputed_100pg',
               'DIA_raw_1ng', 'DIA_imputed_1ng', 'DDA_raw_1ng', 'DDA_imputed_1ng',
               'DIA_raw_10ng', 'DIA_imputed_10ng', 'DDA_raw_10ng', 'DDA_imputed_10ng'),
  'Protein_number' =c(
    1211,2467,457,1030,
    1791,3154,855,1515,
    5264,7061,3095,4676,
    8812,10454, 5061,7794)
)

df_plot <- df_bar %>%
  mutate(
    Mode = ifelse(str_detect(Method, "^DIA"), "DIA", "DDA"),
    Status = ifelse(str_detect(Method, "imputed"), "Imputed", "Raw"),
    Dose = str_extract(Method, "(50pg|100pg|1ng|10ng)")
  )

df_plot$Mode <- factor(df_plot$Mode, levels = c("DIA", "DDA"))
df_plot$Dose <- factor(df_plot$Dose, levels = c('50pg','100pg','1ng', '10ng'))

p <- ggplot(df_plot, aes(x = Mode, y = Protein_number, fill = Status)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  facet_wrap(~ Dose, nrow = 1, scales = "fixed") +
  scale_fill_manual(values = c("Raw" = "#4DBBD5FF", "Imputed" = "#E64B35FF")) +
  labs(
    x = "",
    y = "Protein Count",
    fill = ""
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 13, face = "bold"),
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.7),
    legend.position = "right"
  )

p

ggsave(
  filename = "./figure/Protein_inclusion_numbers.tiff",  
  plot = p,                      
  device = "tiff",                    
  dpi = 300,                          
  width = 7,                         
  height = 4.2,                       
  units = "in"                       
)
```


```{r}
results_plot <- list()
for (method in impute_methods) {
  for (dose in doses) {
    title_prefix <- paste0('./plots_DDA/', method, "-", dose, '.Rdata')
    load(file = title_prefix)
    plot_prefix <- paste0(method, "-", dose)
    results_plot[[plot_prefix]] <- p
  }
}
results_plot_raw <- list()
for (dose in doses) {
    title_prefix <- paste0('./plots_DDA/', "Raw_", dose, '.Rdata')
    load(file = title_prefix)
    plot_prefix <- paste0("Raw_", dose)
    results_plot_raw[[plot_prefix]] <- p
}
```

```{r}
pDDA_Species_Accuracy <- (results_plot_raw[[1]] | results_plot[[1]] | results_plot[[5]] | results_plot[[9]] | results_plot[[13]] | results_plot[[17]]) / (results_plot_raw[[2]] | results_plot[[2]] | results_plot[[6]] | results_plot[[10]] | results_plot[[14]] | results_plot[[18]]) / (results_plot_raw[[3]] | results_plot[[3]] | results_plot[[7]] | results_plot[[11]] | results_plot[[15]] | results_plot[[19]]) / (results_plot_raw[[4]] | results_plot[[4]] | results_plot[[8]] | results_plot[[12]] | results_plot[[16]] | results_plot[[20]]) +   
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') &  
  theme(
    plot.tag = element_text(face = "bold", size = 14),
    plot.tag.position = c(0.02, 0.98)         
  )+
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 18, face = "bold"),
    legend.text      = element_text(size = 16),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold")
  )

ggsave(
  filename = "./figure/DDA_Species_Accuracy.tiff",  
  plot = pDDA_Species_Accuracy,                      
  device = "tiff",                    
  dpi = 300,                         
  width = 27,                         
  height = 20,                       
  units = "in"                        
)
```

# RF, minProb and SoftHybrid
```{r}
pDDA_ThreeMethods_Accuracy <- ((results_plot_raw[[2]] + labs(subtitle = "Missing rate = 18%"))| results_plot[[18]] | results_plot[[2]] | results_plot[[6]] | results_plot[[10]] |results_plot[[14]])  +
    plot_layout(guides = "collect") +
    plot_annotation(tag_levels = 'A') &  
    theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 14, face = "bold"),
    legend.text      = element_text(size = 13),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pDDA_ThreeMethods_Accuracy

ggsave(
  filename = "./figure/DDA_ThreeMethods_Accuracy.tiff",  
  plot = pDDA_ThreeMethods_Accuracy,                      
  device = "tiff",                    
  dpi = 300,                         
  width = 22,                         
  height = 4.3,                     
  units = "in"                       
)
```

```{r}
load(file = './data/results_list_DDA.Rdata')
load(file = './data/results_list_raw_DDA.Rdata')

ground_truth <- c("Ecoli" = -2, "Human" = 0, "Yeast" = 1)
pointwise_error_df <- data.frame()
for (name in names(results_list)) {
  df <- results_list[[name]]
  df <- df %>% filter(Species %in% names(ground_truth))
  df$Truth <- ground_truth[df$Species]
  df$Abs_Error <- abs(df$Log2_Ratio - df$Truth)
  df$Squared_Error <- (df$Log2_Ratio - df$Truth)^2
  df$Method <- name
  pointwise_error_df <- rbind(pointwise_error_df, df)
}

methodwise_error_summary <- pointwise_error_df %>%
  group_by(Method) %>%
  summarise(
    N = n(),
    MAE = mean(Abs_Error, na.rm = TRUE),
    RMSE = sqrt(mean(Squared_Error, na.rm = TRUE)),
    MAE_sd = sd(Abs_Error, na.rm = TRUE),
    se = MAE_sd/sqrt(N),
    lo = MAE - 1.96*se, hi = MAE + 1.96*se, 
    .groups="drop"
  ) %>%
  arrange(MAE)

print(methodwise_error_summary)

specieswise_error_summary <- pointwise_error_df %>%
  group_by(Method, Species) %>%
  summarise(
    N = n(),
    MAE = mean(Abs_Error, na.rm = TRUE),
    RMSE = sqrt(mean(Squared_Error, na.rm = TRUE)),
    MAE_sd = sd(Abs_Error, na.rm = TRUE),
    MAE_sd = sd(Abs_Error, na.rm = TRUE),
    se = MAE_sd/sqrt(N),
    lo = MAE - 1.96*se, hi = MAE + 1.96*se, 
    .groups="drop"
  ) %>%
  arrange(MAE)

print(specieswise_error_summary)
```

### Overall Absolute Error Box Plot
```{r}
plot_df <- pointwise_error_df %>%
  dplyr::select(Protein, Method, Abs_Error) %>%
  mutate(
    Absolute_Error = Abs_Error,
    Dose = str_extract(Method, "50pg|100pg|1ng|10ng"),
    Method_simple = str_remove(Method, "-(50pg|100pg|1ng|10ng)$")
  )

plot_df$Dose <- factor(plot_df$Dose, levels = c("50pg", "100pg", "1ng", "10ng"))
plot_df$Method_simple <- factor(
  plot_df$Method_simple,
  levels = c("Gaussian", "KNN", "minProb", "RF", "SoftHybrid")
)
my_colors <- c(
  "Gaussian"   = "#E69F00",
  "KNN"        = "#56B4E9",
  "minProb"    = "#D55E00",
  "RF"         = "#009E73",
  "SoftHybrid" = "#9966CC"
)
```

```{r}
paired_df_abs <- plot_df %>%
  dplyr::select(Protein, Method, Absolute_Error) %>%
  pivot_wider(
    names_from = Method,
    values_from = Absolute_Error
  )

methods <- c("Gaussian", "KNN", "minProb", "RF")
doses   <- c("50pg", "100pg", "1ng", "10ng")

results_test_abs <- data.frame()

for (d in doses) {
  
  soft_col <- paste0("SoftHybrid-", d)
  
  for (m in methods) {
    
    method_col <- paste0(m, "-", d)
    
    df <- paired_df_abs[, c(soft_col, method_col)]
    df <- na.omit(df)
    
    test <- wilcox.test(
      df[[soft_col]],
      df[[method_col]],
      paired = TRUE,
      alternative = "two.sided"
    )
    
    results_test_abs <- rbind(
      results_test_abs,
      data.frame(
        Dose = d,
        group1 = "SoftHybrid",
        group2 = m,
        N = nrow(df),
        mean_group1 = mean(df[[soft_col]], na.rm = TRUE),
        mean_group2 = mean(df[[method_col]], na.rm = TRUE),
        p_value = test$p.value
      )
    )
  }
}

results_test_abs <- results_test_abs %>%
  group_by(Dose) %>%
  mutate(p_adj = p.adjust(p_value, method = "BH")) %>%
  ungroup() %>%
  mutate(
    signif = case_when(
      p_adj <= 0.001 ~ "***",
      p_adj <= 0.01  ~ "**",
      p_adj <= 0.05  ~ "*",
      TRUE           ~ "ns"
    )
  )

print(results_test_abs)
```

```{r}
y_pos_abs <- plot_df %>%
  group_by(Dose) %>%
  summarise(ymax = max(Absolute_Error, na.rm = TRUE), .groups = "drop")

sig_df_abs <- results_test_abs %>%
  left_join(y_pos_abs, by = "Dose") %>%
  mutate(
    y.position = case_when(
      group2 == "Gaussian" ~ 1.25 * 1.08,
      group2 == "KNN"      ~ 1.25 * 1.20,
      group2 == "minProb"  ~ 1.25 * 1.32,
      group2 == "RF"       ~ 1.25 * 1.44
    )
  )


sig_df_abs$Dose <- factor(sig_df_abs$Dose, levels =  c("50pg", "100pg", "1ng", "10ng"))
```

```{r, warning=FALSE}
pAE_DDA_box <- ggplot(
  plot_df,
  aes(x = Method_simple, y = Absolute_Error, fill = Method_simple)
) +  
  geom_boxplot(
    width = 0.6,
    outlier.shape = NA,
    color = "black",
    linewidth = 0.5,
    fatten = 0
  ) +
  stat_summary(
    fun = mean,
    geom = "point",
    shape = 21,
    size = 2.8,
    fill = "white",
    color = "black",
    stroke = 0.8
  ) +
  facet_wrap(~Dose, ncol = 4, scales = "free_y") +
  coord_cartesian(ylim = c(0, 1.8)) +
  scale_fill_manual(
    name = "Method",        
    values = my_colors
  ) +
  stat_pvalue_manual(
    sig_df_abs,
    label = "signif",
    xmin = "group1",
    xmax = "group2",
    y.position = "y.position",
    tip.length = 0.01,
    bracket.size = 0.5,
    size = 5,
    inherit.aes = FALSE
  ) +
  theme_pub(base_size = 12) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold", colour = "black"),
    axis.text.x      = element_blank(),
    axis.ticks.length= unit(2, "mm"),
    axis.ticks.x     = element_blank(),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    strip.text       = element_text(size = 13, face = "bold") 
  ) +
  labs(
    x = NULL,
    y = "Absolute Error",
    title = "DDA Absolute Error across Imputation Methods"
  )
pAE_DDA_box

ggsave(
  filename = "./figure/AE_DDA_with_significance.tiff",
  plot = pAE_DDA_box,
  device = "tiff",
  dpi = 300,
  width = 8,
  height = 4,
  units = "in"
)
```
### Species-wise box plot
```{r}
plot_df <- pointwise_error_df %>%
  separate(Method, into = c("Method", "Dose"), sep = "-", remove = FALSE) %>%
  mutate(
    Species = factor(Species, levels = c("Human", "Yeast", "Ecoli")),
    Method  = factor(Method, levels = c("Gaussian","KNN","minProb","RF","SoftHybrid")),
    Dose = factor(Dose, levels = c('50pg', '100pg', '1ng', '10ng'))
  )

median_df <- plot_df %>%
  group_by(Method, Species, Dose) %>%
  summarise(
    median_value = median(Log2_Ratio, na.rm = TRUE),
    .groups = "drop"
  )

p_box <- ggplot(plot_df, aes(x = Method, y = Log2_Ratio, fill = Species)) +
  geom_hline(yintercept = 0,  linetype = "dashed", color = "#1b9e77", linewidth = 0.6) +
  geom_hline(yintercept = 1,  linetype = "dashed", color = "#d95f02", linewidth = 0.6) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "#7570b3", linewidth = 0.6) +
  geom_boxplot(
    position = position_dodge(width = 0.75),
    width = 0.65,
    outlier.shape = NA,
    color = "black",
    linewidth = 0.4
  ) +
  #geom_text(
  #  data = median_df,
  #  aes(
  #    x = Method,
  #    y = median_value + 0.2,
  #    label = round(median_value, 2),
  #    group = Species
  #  ),
  #  position = position_dodge(width = 0.75),
  #  vjust = -0.5,
  #  size = 3.5,
  #  fontface = "bold"
  #) +
  scale_fill_manual(values = c(
    "Human" = "#1b9e77",
    "Yeast" = "#d95f02",
    "Ecoli" = "#7570b3"
  )) +
  coord_cartesian(ylim = c(-3, 2)) +
  facet_wrap(~Dose, nrow = 1) +
  labs(
    x = NULL,
    y = "Log2 Ratio",
    title = "DDA Ratio across Doses and Methods"
  ) +
  theme_pub(base_size = 12) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = "bold", colour = "black"),
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    strip.text       = element_text(size = 13, face = "bold") 
  )
p_box 
```
```{r}
ggsave(
  filename = "./figure/logratio_box_DDA.tiff",
  plot = p_box,
  device = "tiff",
  dpi = 300,
  width = 8,
  height = 4,
  units = "in"
)
```

### Distribution of deviation
```{r}
my_colors <- c(
  "Gaussian"   = "#E69F00",
  "KNN"        = "#56B4E9",
  "minProb"    = "#D55E00",
  "RF"         = "#009E73",
  "SoftHybrid" = "#9966CC"
)

p_deviation <- 
ggplot(plot_df, aes(x = Method, y = Abs_Error, fill = Method)) +
  geom_violin() +
  geom_boxplot(width = 0.2) +
  scale_fill_manual(values = my_colors) +
  ylim(0,6) +
  labs(
    x = NULL,
    title = "DDA Ratio Deviation across Methods"
  ) +
  theme_pub(base_size = 12) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = "bold", colour = "black"),
    axis.ticks.length= unit(2, "mm"),
    axis.ticks.x     = element_blank(),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "none", 
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    strip.text       = element_text(size = 13, face = "bold") 
  )
p_deviation
```
```{r}
ggsave(
  filename = "./figure/logratio_deviation_DDA.tiff",
  plot = p_deviation,
  device = "tiff",
  dpi = 300,
  width = 4.2,
  height = 4,
  units = "in"
)
```

### Limma between Sample A and Sample B
```{r}
method_list <- list(
  RF = rf_results,
  minProb = minprob_results,
  SoftHybrid = softhybrid_results,
  KNN = knn_results,
  Gaussian = gaussian_results,
  metadata = list(
    '50pg' = metadata_50pg,
    '100pg' = metadata_100pg,
    '1ng' = metadata_1ng,
    '10ng' = metadata_10ng
  )
)
```

```{r}
run_limma <- function(imputed_data, metadata, method_name, fc_threshold = 0, pval_threshold = 0.05) {
  design <- model.matrix(~0 + Group, metadata)
  colnames(design) <- c("SampleA", "SampleB")

  fit1 <- lmFit(imputed_data, design = design, method = "robust", maxit = 2000)
  contrast <- makeContrasts(SampleA - SampleB, levels = colnames(design))
  fit2 <- contrasts.fit(fit1, contrast)
  fit3 <- eBayes(fit2)

  modtest <- limma::topTable(fit3, number = Inf, sort.by = "none")
  limma_result <- merge(as.data.frame(imputed_data), as.data.frame(modtest), by = "row.names", all = TRUE)
  rownames(limma_result) <- limma_result$Row.names
  limma_result <- limma_result %>%
    dplyr::select(-Row.names) %>%
    arrange(adj.P.Val) %>%
    mutate(gene = rownames(.))

  sig_res <- limma_result %>%
    filter(adj.P.Val < pval_threshold) %>%
    arrange(adj.P.Val)

  limma_result_upregulated <- sig_res %>% filter(logFC > fc_threshold)
  limma_result_downregulated <- sig_res %>% filter(logFC < fc_threshold)

  return(list(
    method = method_name,
    full_result = limma_result,
    significant_DEGs = sig_res,
    upregulated = limma_result_upregulated,
    downregulated = limma_result_downregulated
  ))
}

run_limma_all_groups <- function(method_list) {
    results_list <- list()
    Method_names <- names(method_list[1:5])
    Dose_names <- names(method_list$metadata)
    for (Method in Method_names) {
      cat("Processing Method:", Method, "\n")
        for (Dose in Dose_names) {
                cat("-Processing Dose:", Dose, "\n")
          imputed_data <- method_list[[Method]][[Dose]][["imputed_data"]]
          metadata <- method_list$metadata[[Dose]]
          method_name <- paste0(Method, '_', Dose)
          result <- run_limma(imputed_data, metadata, method_name, fc_threshold = 0, pval_threshold = 0.05)
          results_list[[method_name]] <- result
        }
    }
    return(results_list)
}
```

```{r}
limma_res_list <- run_limma_all_groups(method_list)
```
```{r}
extract_species <- function(rownms) {
  sp <- str_extract(rownms, "(?<=_)HUMAN$|(?<=_)YEAST$|(?<=_)ECOLI$")
  sp
}

species_to_truthdir <- function(sp) {
  case_when(
    sp == "YEAST" ~  1L,
    sp == "ECOLI" ~ -1L,   
    sp == "HUMAN" ~  0L,
    TRUE ~ NA_integer_
  )
}

call_dir_from_limma <- function(df, p_thresh = 0.05, fc_thresh = 0) {
  case_when(
    df$adj.P.Val < p_thresh & df$logFC >  fc_thresh ~  1L,
    df$adj.P.Val < p_thresh & df$logFC <  fc_thresh ~ -1L,
    TRUE ~ 0L
  )
}
```

```{r}
evaluate_one <- function(limma_result_tbl, method_name) {
  df <- limma_result_tbl$full_result %>%
    dplyr::mutate(
      species   = extract_species(rownames(.)),
      truth_dir = species_to_truthdir(species),
      call_dir  = call_dir_from_limma(., p_thresh = 0.05, fc_thresh = 0)
    ) %>%
    dplyr::filter(!is.na(t), !is.na(truth_dir))

  TP_up <- sum(df$truth_dir ==  1 & df$call_dir ==  1)
  FP_up <- sum(df$truth_dir !=  1 & df$call_dir ==  1)
  FN_up <- sum(df$truth_dir ==  1 & df$call_dir !=  1)
  TN_up <- sum(df$truth_dir !=  1 & df$call_dir !=  1)

  TP_down <- sum(df$truth_dir == -1 & df$call_dir == -1)
  FP_down <- sum(df$truth_dir != -1 & df$call_dir == -1)
  FN_down <- sum(df$truth_dir == -1 & df$call_dir != -1)
  TN_down <- sum(df$truth_dir != -1 & df$call_dir != -1)

  TP_neu <- sum(df$truth_dir ==  0 & df$call_dir ==  0)
  FP_neu <- sum(df$truth_dir !=  0 & df$call_dir ==  0)
  FN_neu <- sum(df$truth_dir ==  0 & df$call_dir !=  0)
  TN_neu <- sum(df$truth_dir !=  0 & df$call_dir !=  0)

  conf_long <- dplyr::bind_rows(
    tibble::tibble(MethodDose = method_name, Task = "Up(yeast)",   TP=TP_up,   FP=FP_up,   FN=FN_up,   TN=TN_up),
    tibble::tibble(MethodDose = method_name, Task = "Down(ecoli)", TP=TP_down, FP=FP_down, FN=FN_down, TN=TN_down),
    tibble::tibble(MethodDose = method_name, Task = "Neutral(human)", TP=TP_neu, FP=FP_neu, FN=FN_neu, TN=TN_neu)
  ) %>%
    dplyr::mutate(
      Accuracy = (TP + TN) / (TP + FP + FN + TN),
      Precision = TP / pmax(TP + FP, 1),
      Recall = TP / pmax(TP + FN, 1),
      F1 = 2 * Precision * Recall / pmax(Precision + Recall, 1e-9)
    )

  roc_up <- pROC::roc(response = (df$truth_dir == 1),
                      predictor = df$t,
                      levels = c(FALSE, TRUE),
                      direction = "<",
                      quiet = TRUE)
  auc_up <- as.numeric(pROC::auc(roc_up))

  roc_down <- pROC::roc(response = (df$truth_dir == -1),
                        predictor = df$t,
                        levels = c(FALSE, TRUE),
                        direction = ">",
                        quiet = TRUE)
  auc_down <- as.numeric(pROC::auc(roc_down))

  roc_neu <- pROC::roc(response = (df$truth_dir == 0),
                       predictor = abs(df$t),
                       levels = c(FALSE, TRUE),
                       direction = ">",
                       quiet = TRUE)
  auc_neu <- as.numeric(pROC::auc(roc_neu))

  macro_auc <- mean(c(auc_up, auc_down, auc_neu))

  roc_df_up <- tibble::tibble(MethodDose = method_name, Task = "Up(yeast)",
                              FPR = 1 - roc_up$specificities,  TPR = roc_up$sensitivities)
  roc_df_down <- tibble::tibble(MethodDose = method_name, Task = "Down(ecoli)",
                                FPR = 1 - roc_down$specificities, TPR = roc_down$sensitivities)
  roc_df_neu <- tibble::tibble(MethodDose = method_name, Task = "Neutral(human)",
                               FPR = 1 - roc_neu$specificities,  TPR = roc_neu$sensitivities)

  auc_tbl <- tibble::tibble(
    MethodDose = method_name,
    Task = c("Up(yeast)", "Down(ecoli)", "Neutral(human)", "Macro"),
    AUC  = c(auc_up, auc_down, auc_neu, macro_auc)
  )

  list(confusion = conf_long,
       roc_curve = dplyr::bind_rows(roc_df_up, roc_df_down, roc_df_neu),
       auc = auc_tbl)
}
```

```{r}
all_evals <- imap(limma_res_list, ~ evaluate_one(.x, .y))

confusion_summary <- bind_rows(lapply(all_evals, `[[`, "confusion"))
roc_curves_df     <- bind_rows(lapply(all_evals, `[[`, "roc_curve"))
auc_summary       <- bind_rows(lapply(all_evals, `[[`, "auc"))

split_md <- function(x) {
  parts <- str_match(x, "^(.*?)_(50pg|100pg|1ng|10ng)$")
  tibble(MethodDose = x, Method = parts[,2], Dose = parts[,3])
}
md_map <- split_md(unique(roc_curves_df$MethodDose))

confusion_summary <- confusion_summary %>% left_join(md_map, by="MethodDose")
roc_curves_df     <- roc_curves_df     %>% left_join(md_map, by="MethodDose")
auc_summary       <- auc_summary       %>% left_join(md_map, by="MethodDose")
```

```{r}
my_colors <- c(
  "Gaussian"   = "#E69F00",
  "KNN"        = "#56B4E9",
  "RF"         = "#009E73",
  "SoftHybrid" = "#9966CC",
  "minProb"    = "#D55E00"
)

dose_lv   <- c("50pg","100pg","1ng","10ng")
method_lv <- c("Gaussian","KNN","minProb","RF","SoftHybrid")
roc_curves_df$Dose <- factor(roc_curves_df$Dose, levels = dose_lv)

pROC <- ggplot(roc_curves_df, aes(x = FPR, y = TPR, color = Method)) +
  geom_path(size = 0.8, alpha = 0.9) +
  scale_color_manual(values = my_colors) +
  facet_grid(Task ~ Dose) +
  coord_equal() +
  theme_bw() +
  labs(title = "ROC Curves by Method and Dose: DDA",
       subtitle = "Up task: Positive=Yeast (score=t); Down task: Positive=E. coli (score=-t)",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_blank(),
    axis.text.y      = element_blank(),
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )

pROC

ggsave(
  filename = "./figure/ROC_DDA.tiff",  
  plot = pROC,                      
  device = "tiff",                    
  dpi = 300,                          
  width = 8,                         
  height = 8,                        
  units = "in"                       
)
```


```{r}
dose_lv   <- c("50pg","100pg","1ng","10ng")
method_lv <- c("Gaussian","KNN","minProb","RF","SoftHybrid")

tasks_keep <- c("Up(yeast)", "Down(ecoli)", "Neutral(human)")

fpr_grid <- seq(0, 1, length.out = 501)

roc_interp <- roc_curves_df %>%
  dplyr::filter(Task %in% tasks_keep) %>%
  dplyr::arrange(MethodDose, Task, FPR) %>%
  dplyr::group_by(MethodDose, Task) %>%
  dplyr::group_modify(function(.x, .y) {
    d <- .x %>% arrange(FPR, TPR)
    x <- c(0, d$FPR, 1)
    y <- c(0, d$TPR, 1)
    tibble(FPR = fpr_grid,
           TPR = approx(x = x, y = y, xout = fpr_grid, rule = 2, ties = "ordered")$y)
  }) %>%
  dplyr::ungroup()

macro_interp <- roc_interp %>%
  dplyr::group_by(MethodDose, FPR) %>%
  dplyr::summarise(TPR_macro = mean(TPR, na.rm = TRUE), .groups = "drop") %>%
  dplyr::left_join(md_map, by = "MethodDose") %>%
  dplyr::mutate(
    Dose   = factor(Dose,   levels = dose_lv),
    Method = factor(Method, levels = method_lv)
  )

p_macroROC <- ggplot(macro_interp, aes(x = FPR, y = TPR_macro, color = Method)) +
  geom_path(size = 0.9, alpha = 0.95) +
  scale_color_manual(values = my_colors) +
  facet_grid(. ~ Dose) +
  coord_equal() +
  theme_bw() +
  labs(
    title = "Macro ROC Curves by Method and Dose",
    subtitle = "Macro = average TPR over Up(yeast), Down(ecoli), Neutral(human) at matched FPR",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (macro-averaged)"
  ) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold", colour = "black"),
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )

p_macroROC

ggsave(
  filename = "./figure/macroROC_DDA.tiff",  
  plot = p_macroROC,                      
  device = "tiff",                    
  dpi = 300,                         
  width = 12,                         
  height = 7,                        
  units = "in"                       
)
```
```{r}
macro_auc <- macro_interp %>%
  arrange(MethodDose, FPR) %>%
  group_by(MethodDose, Dose, Method) %>%
  summarise(
    AUC = sum(diff(FPR) * zoo::rollmean(TPR_macro, 2), na.rm = TRUE),
    .groups = "drop"
  )
best_by_dose <- macro_auc %>%
  group_by(Dose) %>%
  slice_max(AUC, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(is_best = TRUE) %>%
  dplyr::select(Dose, Method, is_best)

plot_df <- macro_interp %>%
  left_join(best_by_dose, by = c("Dose","Method")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

label_pos <- plot_df %>%
  group_by(Dose, Method) %>%
  arrange(FPR, .by_group = TRUE) %>%
  summarise(
    TPR_macro = approx(
      x = FPR,
      y = TPR_macro,
      xout = 0.20,
      rule = 2,
      ties = "ordered"
    )$y %>% as.numeric(),
    .groups = "drop"
  ) %>%
  mutate(FPR = 0.20) %>%
  left_join(macro_auc, by = c("Dose","Method")) %>%
  mutate(label = sprintf("%s: AUC=%.3f", as.character(Method), AUC))

linetypes <- c(
  Gaussian   = "solid",
  KNN        = "dashed",
  minProb    = "dotdash",
  RF         = "twodash",
  SoftHybrid = "solid"
)

labels_under_best <- macro_auc %>%
  group_by(Dose) %>%
  arrange(desc(AUC), .by_group = TRUE) %>%
  mutate(
    rank = row_number(),
    x = 0.012,                        
    y = 0.985 - 0.01 * rank,            
    label = sprintf("%s: AUC=%.3f", as.character(Method), AUC)
  ) %>%
  ungroup()

p_macroROC_zoom <-
  ggplot(plot_df, aes(x = FPR, y = TPR_macro, color = Method, linetype = Method)) +
  geom_path(aes(size = ifelse(is_best, 1.2, 0.8)), alpha = 0.95) +
  scale_size_identity() +
  scale_color_manual(values = my_colors) +
  scale_linetype_manual(values = linetypes) +
  facet_grid(. ~ Dose) +
  coord_cartesian(xlim = c(0, 0.3), ylim = c(0.80, 1.00), expand = FALSE) +
  theme_bw() +
  labs(
    title = "Macro ROC (Zoomed) by Method and Dose: DDA",
    subtitle = "Macro = mean TPR over Up(yeast), Down(ecoli), Neutral(human) at matched FPR",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (macro-averaged)"
  ) +
  theme(
    text = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    legend.position = "right"
  ) +
  geom_text(
    data = best_by_dose,
    inherit.aes = FALSE,
    aes(x = 0.005, y = 0.995, label = paste0("★ Best: ", Method)),
    hjust = 0, vjust = 1, size = 3.6, fontface = "bold"
  ) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold", colour = "black"),
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    strip.text       = element_text(size = 13, face = "bold") 
  )

p_macroROC_zoom

ggsave(
  filename = "./figure/macroROC_zoom_DDA.tiff", 
  plot = p_macroROC_zoom,                     
  device = "tiff",                   
  dpi = 300,                         
  width = 12,                       
  height = 4,                     
  units = "in"                        
)
```

```{r}
df_numline <- auc_summary %>%
  filter(Task == "Macro") %>%
  dplyr::select(MethodDose, AUC) %>%
  left_join(md_map, by = "MethodDose") %>%
  mutate(
    Dose   = factor(Dose, levels = c("50pg","100pg","1ng","10ng")),
    Method = factor(Method, levels = c("Gaussian","KNN","minProb","RF","SoftHybrid"))
  )

shape_vec <- c(
  "Gaussian"   = 15,
  "KNN"        = 3,
  "minProb"    = 18,
  "RF"         = 17,
  "SoftHybrid" = 16
)

x_min <- 0.75

p_numline <- ggplot(
  df_numline,
  aes(x = AUC, y = Method, color = Method, shape = Method)
) +
  geom_segment(
    aes(x = x_min, xend = AUC, y = Method, yend = Method),
    linewidth = 0.8,
    alpha = 0.8
  ) +
  geom_point(size = 3.5, stroke = 1) +
  facet_wrap(~Dose, nrow = 1) +   # ← 横向排
  scale_color_manual(values = my_colors) +
  scale_shape_manual(values = shape_vec) +
  scale_x_continuous(
  limits = c(0.75, 1),
  breaks = seq(0.75, 1, 0.1)
) +
  labs(x = "Mean-macro AUC", y = NULL) +
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.spacing.x    = unit(10, "pt"),
    legend.position    = "right",
    panel.border       = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text               = element_text(size = 13),
    axis.text          = element_text(size = 13, face = "bold", colour = "black"),
    axis.title         = element_text(size = 14, face = "bold"),
    legend.title       = element_text(size = 13, face = "bold"),
    legend.text        = element_text(size = 12),
    strip.text         = element_text(size = 13, face = "bold")
  )

p_numline
```
```{r}
ggsave("./figure/mean_macro_numline_clean_DDA.tiff", 
       p_numline, 
       width = 12, 
       height = 2, 
       dpi = 300, 
       device = "tiff")
```

### DeLong Significance Test
```{r}
build_roc_input_one <- function(limma_result_tbl, method_name) {
  
  parts <- str_match(method_name, "^(.*?)_(50pg|100pg|1ng|10ng)$")
  method <- parts[,2]
  dose   <- parts[,3]
  
  df <- limma_result_tbl$full_result %>%
    tibble::rownames_to_column("Protein") %>%
    mutate(
      species   = extract_species(Protein),
      truth_dir = species_to_truthdir(species)
    ) %>%
    filter(!is.na(.data$t), !is.na(truth_dir)) %>%
    dplyr::select(Protein, t, truth_dir)
  
  bind_rows(
    df %>% transmute(Protein, MethodDose = method_name, Method = method, Dose = dose,
                     Task = "Up(yeast)", y_true = truth_dir == 1, score = t),
    df %>% transmute(Protein, MethodDose = method_name, Method = method, Dose = dose,
                     Task = "Down(ecoli)", y_true = truth_dir == -1, score = t),
    df %>% transmute(Protein, MethodDose = method_name, Method = method, Dose = dose,
                     Task = "Neutral(human)", y_true = truth_dir == 0, score = abs(t))
  )
}

roc_input_df <- imap_dfr(limma_res_list, build_roc_input_one)

safe_delong_test <- function(df_sub, method1, method2, task_name) {
  
  x1 <- df_sub %>% filter(Method == method1) %>% arrange(Protein)
  x2 <- df_sub %>% filter(Method == method2) %>% arrange(Protein)
  
  common_protein <- intersect(x1$Protein, x2$Protein)
  x1 <- x1 %>% filter(Protein %in% common_protein) %>% arrange(Protein)
  x2 <- x2 %>% filter(Protein %in% common_protein) %>% arrange(Protein)
  
  if (nrow(x1) == 0 || nrow(x2) == 0) {
    return(tibble(Method1 = method1, Method2 = method2, n = 0,
                  AUC1 = NA_real_, AUC2 = NA_real_, p_value = NA_real_))
  }
  
  if (!identical(x1$Protein, x2$Protein)) stop("Protein mismatch")
  if (!identical(x1$y_true, x2$y_true)) stop("y_true mismatch")
  
  y_true <- x1$y_true
  
  if (sum(y_true) == 0 || sum(!y_true) == 0) {
    return(tibble(Method1 = method1, Method2 = method2, n = length(y_true),
                  AUC1 = NA_real_, AUC2 = NA_real_, p_value = NA_real_))
  }
  
  direction_use <- case_when(
    task_name == "Up(yeast)" ~ "<",
    task_name == "Down(ecoli)" ~ ">",
    task_name == "Neutral(human)" ~ ">"
  )
  
  roc1 <- pROC::roc(y_true, x1$score, levels = c(FALSE, TRUE),
                    direction = direction_use, quiet = TRUE)
  roc2 <- pROC::roc(y_true, x2$score, levels = c(FALSE, TRUE),
                    direction = direction_use, quiet = TRUE)
  
  tt <- pROC::roc.test(roc1, roc2, method = "delong", paired = TRUE)
  
  tibble(
    Method1 = method1,
    Method2 = method2,
    n = length(y_true),
    AUC1 = as.numeric(pROC::auc(roc1)),
    AUC2 = as.numeric(pROC::auc(roc2)),
    p_value = tt$p.value
  )
}

methods_lv <- c("Gaussian","KNN","minProb","RF","SoftHybrid")
ref_method <- "SoftHybrid"
other_methods <- setdiff(methods_lv, ref_method)

delong_soft_vs_others <- roc_input_df %>%
  group_by(Dose, Task) %>%
  group_modify(~{
    task_now <- .y$Task[[1]]
    
    bind_rows(lapply(other_methods, function(m) {
      safe_delong_test(.x, ref_method, m, task_now)
    }))
  }) %>%
  ungroup() %>%
  group_by(Dose, Task) %>%
  mutate(p_adj_BH = p.adjust(p_value, method = "BH")) %>%
  ungroup()

delong_soft_vs_others

## Save significance result table
write.csv(
  delong_soft_vs_others,
  file = "./output_DDA/delong_soft_vs_others_DDA.csv",
  row.names = FALSE
)
```

```{r}
sessionInfo()
```
