---
title: "3_ImputationBenchmarking"
author: "Yixin Shi"
date: "2025-06-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(magrittr, ggplot2, ggVennDiagram, pheatmap, ggridges, umap, limma, dplyr, tidyr, viridis, gridExtra, stringr, ggsci, clusterProfiler, org.Mm.eg.db, UpSetR, missForest,ComplexHeatmap, WGCNA, org.Hs.eg.db, GOSemSim, ggrepel, VennDiagram, GSVA, tibble, purrr, scales, imputeLCMD, patchwork, mclust, cluster, ggpubr)
```

# Data Preperation
## Read Data
```{r, warning=FALSE}
rm(list = ls())
getwd()

intensity_bulk <- read.csv("./data/report.unique_genes_matrix_50cell_n_rerun.tsv", sep = "\t", header = T)
rownames(intensity_bulk) <- intensity_bulk$Genes
intensity_bulk <- intensity_bulk[,-c(1:3)]
sample <- unlist(lapply(strsplit(colnames(intensity_bulk), split = '_', fixed = TRUE), function(x) x[12]))
colnames(intensity_bulk) <- unlist(lapply(strsplit(sample, split = '.', fixed = TRUE), function(x) x[1]))
demo <- intensity_bulk[,5:8]
intensity_bulk <- cbind(demo, intensity_bulk[,-c(5:8)])
```

```{r}
intensity_scp <- read.csv("./data/report.unique_genes_matrix_1cell_n_rerun.tsv", sep = "\t", header = T)
rownames(intensity_scp) <- intensity_scp$Genes
intensity_scp <- intensity_scp[,-c(1:3)]
sample <- unlist(lapply(strsplit(colnames(intensity_scp), split = '_', fixed = TRUE), function(x) x[12]))
group <- unlist(lapply(strsplit(colnames(intensity_scp), split = '_', fixed = TRUE), function(x) x[6]))
colnames(intensity_scp) <- unlist(lapply(strsplit(sample, split = '.', fixed = TRUE), function(x) x[1]))

demoGSC <- intensity_scp[,21:40]
demoAST <- intensity_scp[,1:20]
intensity_scp <- cbind(demoGSC,demoAST,intensity_scp[,-(1:40)])
```

## Prepare Metadata File
```{r, warning=FALSE}
### for annotating sample information, at least two columns, one for sample name and one for grouping information
metadata_bulk <- data.frame("sample" = colnames(intensity_bulk) %>%
                        as.factor(),"group" = rep(c("GSC", 'AST', "MAC", "MIC"), each =4))
metadata_bulk$group <- metadata_bulk$group %>%
                                factor(levels = c("GSC", 'AST', "MAC", "MIC"))
### mark the number of protein types detected
metadata_bulk$ID <- apply(intensity_bulk, 2, function(x) sum(!is.na(x)))
rownames(metadata_bulk) <- metadata_bulk$sample
```

```{r}
metadata_scp <- data.frame("sample" = colnames(intensity_scp) %>%
                        as.factor(),"group" = rep(c("GSC", 'AST', "MAC", "MIC"), each = 20))
metadata_scp$group <- metadata_scp$group %>%
                                factor(levels = c("GSC", 'AST', "MAC", "MIC"))
### mark the number of protein types detected
metadata_scp$ID <- apply(intensity_scp, 2, function(x) sum(!is.na(x)))
rownames(metadata_scp) <- metadata_scp$sample
```

## Data matrix manipulation
```{r}
intensity_bulk_long <- intensity_bulk %>% 
  mutate(gene = rownames(.)) %>%
  pivot_longer(cols = colnames(intensity_bulk), 
               names_to = "sample", 
               values_to = "intensity")
intensity_bulk_long$group <- plyr::mapvalues(intensity_bulk_long$sample,
                                       from = metadata_bulk$sample,
                                       to = metadata_bulk$group)
intensity_bulk_long$group <- factor(intensity_bulk_long$group, levels = c(1,2,3,4))
levels(intensity_bulk_long$group) <- c("GSC", 'AST', "MAC", "MIC")
intensity_bulk_long$sample <- factor(intensity_bulk_long$sample, levels = metadata_bulk$sample)
```

```{r}
intensity_scp_long <- intensity_scp %>% 
  mutate(gene = rownames(.)) %>%
  pivot_longer(cols = colnames(intensity_scp), 
               names_to = "sample", 
               values_to = "intensity")
intensity_scp_long$group <- plyr::mapvalues(intensity_scp_long$sample,
                                       from = metadata_scp$sample,
                                       to = metadata_scp$group)
intensity_scp_long$group <- factor(intensity_scp_long$group, levels = c(1,2,3,4))
levels(intensity_scp_long$group) <- c("GSC", 'AST', "MAC", "MIC")
intensity_scp_long$sample <- factor(intensity_scp_long$sample, levels = metadata_scp$sample)
```

## Correlation of abundance and missing rate : The proteins with high missing rates are generally less abundant (MNAR)
```{r}
tb_cor = data.frame(row.names = row.names(intensity_bulk), 
                    missing_rate = rowSums(is.na(intensity_bulk)) / ncol(intensity_bulk),
                    abundance = rowMeans(log2(intensity_bulk), na.rm = T))

ggplot(tb_cor, aes(x = factor(missing_rate), y = abundance)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(size = 0.7) +
  # geom_smooth(method = "lm", se = F) +
  theme_bw() +
  xlab("missing_rate") +
  ylab("Abundance") +
  ggtitle("Correlation of abundance and missing number")

print(with(tb_cor, cor.test(missing_rate, abundance, method = "spearman")))
```

## Filter expression profile in bulk
```{r}
### Manipulate data by log2
df_bulk_log2 <- log2(intensity_bulk)
df_bulk_long_log2 <- intensity_bulk_long
df_bulk_long_log2$intensity <- log2(df_bulk_long_log2$intensity)

geneName1 <- tb_cor %>%
  filter(missing_rate < 0.9) %>%
  rownames(.)

df_bulk_long_filter <- df_bulk_long_log2 %>%
  filter(gene %in% geneName1)

df_bulk_filter <- df_bulk_long_filter [,-4] %>%
  pivot_wider(names_from = sample,
              values_from = intensity) %>%
  as.data.frame()
rownames(df_bulk_filter )<- df_bulk_filter$gene
df_bulk_filter <- df_bulk_filter[,-1]
```

## Normalisation of ground truth
```{r}
### Data normalization function
median_centering <- function(df) {
  df_norm <- as.data.frame(apply(df, 2, function(x) {
                        gMedian = median(x, na.rm = TRUE)
                        x - gMedian
  }
    ))
  df_long_norm <- df_norm %>%
    mutate(gene = rownames(.)) %>%
    pivot_longer(cols = colnames(df_norm),
                 names_to = 'sample',
                 values_to = 'intensity')
  df_long_norm$group <- plyr::mapvalues(df_long_norm$sample,
                                        from = metadata_bulk$sample,
                                        to = metadata_bulk$group)
  df_long_norm$sample <- factor(df_long_norm$sample, levels = metadata_bulk$sample)
  
  return(df_long_norm)
}
### Normalize data
df_bulk_long_norm <- median_centering(df_bulk_filter)
df_bulk_long_norm$group <- factor(df_bulk_long_norm$group, levels = c(1,2,3,4))
levels(df_bulk_long_norm$group) <- c("GSC", 'AST', "MAC", "MIC")
df_bulk_norm <- df_bulk_long_norm %>% 
  dplyr::select(-group) %>%
  pivot_wider(names_from = sample, values_from = intensity) %>% 
  as.data.frame()
rownames(df_bulk_norm) <- df_bulk_norm$gene
df_bulk_norm <- df_bulk_norm[,-1]
```

## Correlation of abundance and missing rate : The proteins with high missing rates are generally less abundant (MNAR)
```{r}
tb_cor = data.frame(row.names = row.names(intensity_scp), 
                    missing_rate = rowSums(is.na(intensity_scp)) / ncol(intensity_scp),
                    abundance = rowMeans(log2(intensity_scp), na.rm = T))

ggplot(tb_cor, aes(x = factor(missing_rate), y = abundance)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(size = 0.7) +
  # geom_smooth(method = "lm", se = F) +
  theme_bw() +
  xlab("missing_rate") +
  ylab("Abundance") +
  ggtitle("Correlation of abundance and missing number")

print(with(tb_cor, cor.test(missing_rate, abundance, method = "spearman")))
```

## Filter expression profile in scp: sample with ID <1000
```{r}
### Manipulate data by log2
df_scp_log2 <- log2(intensity_scp)
df_scp_long_log2 <- intensity_scp_long
df_scp_long_log2$intensity <- log2(df_scp_long_log2$intensity)

metadata_scp$sample <- as.character(metadata_scp$sample)
filter_sample <- metadata_scp[metadata_scp$ID >= 1000, "sample"]
metadata_scp <- metadata_scp[filter_sample,]

geneName2 <- tb_cor %>%
  filter(missing_rate < 0.9) %>%
  rownames(.)

df_scp_long_filter <- df_scp_long_log2 %>%
  filter(gene %in% geneName2) %>%
  filter(sample %in% filter_sample)

df_scp_filter <- df_scp_long_filter[,-4] %>%
  pivot_wider(names_from = sample,
              values_from = intensity) %>%
  as.data.frame()
rownames(df_scp_filter )<- df_scp_filter$gene
df_scp_filter <- df_scp_filter[,-1]
```

## Normalisation of scp matrix
```{r}
### Data normalization function
median_centering <- function(df) {
  df_norm <- as.data.frame(apply(df, 2, function(x) {
                        gMedian = median(x, na.rm = TRUE)
                        x - gMedian
  }
    ))
  df_long_norm <- df_norm %>%
    mutate(gene = rownames(.)) %>%
    pivot_longer(cols = colnames(df_norm),
                 names_to = 'sample',
                 values_to = 'intensity')
  df_long_norm$group <- plyr::mapvalues(df_long_norm$sample,
                                        from = metadata_scp$sample,
                                        to = metadata_scp$group)
  df_long_norm$sample <- factor(df_long_norm$sample, levels = metadata_scp$sample)
  
  return(df_long_norm)
}
### Normalize data
df_scp_long_norm <- median_centering(df_scp_filter)
df_scp_long_norm$group <- factor(df_scp_long_norm$group, levels = c(1,2,3,4))
levels(df_scp_long_norm$group) <- c("GSC", 'AST', "MAC", "MIC")
df_scp_norm <- df_scp_long_norm %>% 
  dplyr::select(-group) %>%
  pivot_wider(names_from = sample, values_from = intensity) %>% 
  as.data.frame()
rownames(df_scp_norm) <- df_scp_norm$gene
df_scp_norm <- df_scp_norm[,-1]
```

## Find the intersected proteins between SCP and bulk
```{r}
protein_scp <- rownames(df_scp_norm)
protein_bulk <- rownames(df_bulk_norm)
common_protein <- intersect(protein_bulk, protein_scp)

df_scp <- df_scp_norm[common_protein,]
df_bulk <- df_bulk_norm[common_protein,]

save(df_bulk, file = './data/bulk_benchmarking.Rdata')
save(metadata_bulk, file = './data/metadata_bulk_benchmarking.Rdata')
```

# Categorize NA: class A-cell type specific protein; Class B-universal protein
```{r}

## Marker Candidate protein: in at least one group, >=75% replicates are NA values and in at least one group, >=75% replicates have non-missing values
df_stats <- df_bulk_long_norm %>%
  group_by(gene, group) %>%
  summarise(
    NA_rate    = mean(is.na(intensity)),
    NonNA_rate = 1 - NA_rate,
    MeanExpr   = mean(intensity, na.rm = TRUE),
    .groups = "drop"
  )

marker_candidate <- vector("character")
for (prot in common_protein) {
  prot_df <- df_stats %>% filter(gene == prot)
  cond1 <- any(prot_df$NA_rate > 0.75)
  cond2 <- any(prot_df$NonNA_rate > 0.75)
  if (cond1 && cond2) {
    marker_candidate <- c(marker_candidate, prot)
  }
}

# 最终结果
length(marker_candidate)
head(marker_candidate)

marker_others <- df_bulk %>%
  filter(!rownames(.) %in% marker_candidate) %>%
  rownames(.)
```
# Imputation Method I: Random Forest of ALL proteins
```{r}
#set.seed(123)

#time_taken_rf <- system.time({
#  res_all  <- missForest(as.data.frame(df_scp), verbose=FALSE)
#  imp_all  <- res_all$ximp
#})

#save(imp_all, file = './data/imp_all_benchmarking.Rdata')
load(file = './data/imp_all_benchmarking.Rdata')
#print(time_taken_rf)
```

# Imputation Method II: Minimal Probability of ALL proteins
```{r}
#set.seed(25)

#time_taken_minProb <- system.time({
#  minProb_all <- impute.MinProb(
#  df_scp,
#  q = 0.01,
#  tune.sigma = 1
#)
#})

#save(minProb_all, file = './data/minProb_all_benchmarking.Rdata')
load(file = './data/minProb_all_benchmarking.Rdata')
#print(time_taken_minProb)
```
# Imputation Method III: KNN of ALL proteins
```{r}
#set.seed(25)

#time_taken_knn <- system.time({
#  knn_impute <- impute.knn(as.matrix(df_scp), k = 10)
#  knn_all <- as.data.frame(knn_impute[[1]])
#})

#save(knn_all, file = './data/knn_all_benchmarking.Rdata')
load(file = './data/knn_all_benchmarking.Rdata')
#print(time_taken_knn)
```
# Imputation Method IV: Gaussian sample of ALL proteins
```{r}
#set.seed(25)
#gaussian_impute <- function(df, width = 0.3, downshift = 1.8) {
#  imputed_df <- df
#  all_values <- unlist(df)
#  observed_values <- all_values[!is.na(all_values)]
  
#  mu <- mean(observed_values, na.rm = TRUE)
#  sigma <- sd(observed_values, na.rm = TRUE)
  
#  impute_mean <- mu - downshift * sigma
#  impute_sd <- sigma * width
  
#  for (i in 1:nrow(df)) {
#    for (j in 1:ncol(df)) {
#      if (is.na(df[i, j])) {
#        imputed_df[i, j] <- rnorm(1, mean = impute_mean, sd = impute_sd)
#      }
#    }
#  }
#  return(imputed_df)
#}

#time_taken_GS <- system.time({
#  GS_all <- gaussian_impute(df_scp)
#})

#save(GS_all, file = './data/GS_All_benchmarking.Rdata')
load(file = './data/GS_All_benchmarking.Rdata')
#print(time_taken_GS)
```
# Imputation Method V: Soft Hybrid Imputation
## Assessment of missing rate
```{r}
miss_cor <- data.frame(
  missing_rate    = rowSums(is.na(df_scp)) / ncol(df_scp),
  mean_abundance  = rowMeans(df_scp, na.rm = TRUE)
)
ggplot(miss_cor, aes(x = missing_rate, y = mean_abundance)) +
  geom_point(size = 0.7, alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  theme_bw() +
  xlab("Missing Rate") +
  ylab("Mean log2 Abundance") +
  ggtitle("Correlation of Abundance and Missing Rate")
```
```{r}
library(ggplot2)
library(scales)

loess_fit <- loess(mean_abundance ~ missing_rate, data = miss_cor, span = 0.75)
x_vals <- seq(min(miss_cor$missing_rate), max(miss_cor$missing_rate), length.out = 1000)
y_vals <- predict(loess_fit, newdata = data.frame(missing_rate = x_vals))
x_start <- x_vals[1]; y_start <- y_vals[1]
x_end   <- x_vals[length(x_vals)]; y_end <- y_vals[length(y_vals)]
line_vec <- c(x_end - x_start, y_end - y_start)
distances <- sapply(1:length(x_vals), function(i) {
  point_vec <- c(x_vals[i] - x_start, y_vals[i] - y_start)
  cross_prod <- abs(line_vec[1] * point_vec[2] - line_vec[2] * point_vec[1])
  norm_line <- sqrt(sum(line_vec^2))
  return(cross_prod / norm_line)
})
max_index <- which.max(distances)
elbow_x <- x_vals[max_index]
elbow_y <- y_vals[max_index]

ggplot(miss_cor, aes(x = missing_rate, y = mean_abundance)) +
  geom_point(size = 0.7, alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.75, color = "blue") +
  geom_abline(
    slope = (y_end - y_start) / (x_end - x_start),
    intercept = y_start - ((y_end - y_start) / (x_end - x_start)) * x_start,
    color = "gray50", linetype = "dashed"
  ) +
  geom_point(aes(x = elbow_x, y = elbow_y), color = "red", size = 2) +
  geom_text(aes(x = elbow_x + 0.2,
                y = elbow_y + 5,
                label = paste0("Elbow @ ", percent(elbow_x, accuracy = 1))),
            color = "red") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  theme_bw() +
  xlab("Missing Rate") +
  ylab("Mean log2 Abundance") +
  ggtitle("Loess-Based Elbow Point by Maximum Distance Method")

theme_pub <- function(base_size = 10) {
  ggplot2::theme_classic(base_size = base_size) %+replace%
    ggplot2::theme(
      panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
      axis.title       = element_text(size = base_size + 1, face = "bold"),
      axis.text        = element_text(size = base_size),
      axis.ticks.length= unit(2, "mm"),
      plot.title       = element_text(size = base_size + 2, face = "bold", hjust = 0.5),
      legend.position  = "right",
      legend.title     = element_text(size = base_size),
      legend.text      = element_text(size = base_size - 1)
    )
}

find_missing_rate_elbow <- function(df_norm,
                                    span = 0.75,
                                    n_points = 1000,
                                    point_size = 0.8,
                                    point_alpha = 0.8,
                                    save_path = NULL,   
                                    group_name = NULL,
                                    tiff_width = 4,  
                                    tiff_height = 3.5,
                                    tiff_dpi = 600) {
  
  stopifnot(is.matrix(df_norm) || is.data.frame(df_norm))
  if (!requireNamespace("ggplot2", quietly = TRUE)) stop("Please install 'ggplot2'.")
  if (!requireNamespace("scales", quietly = TRUE))  stop("Please install 'scales'.")
  
  df_norm <- as.matrix(df_norm)
  
  miss_cor <- data.frame(
    missing_rate   = rowSums(is.na(df_norm)) / ncol(df_norm),
    mean_abundance = rowMeans(df_norm, na.rm = TRUE)
  )
  
  fit <- stats::loess(mean_abundance ~ missing_rate, data = miss_cor, span = span)
  
  x_vals <- seq(min(miss_cor$missing_rate, na.rm = TRUE),
                max(miss_cor$missing_rate, na.rm = TRUE),
                length.out = n_points)
  y_vals <- stats::predict(fit, newdata = data.frame(missing_rate = x_vals))
  
  keep   <- !is.na(y_vals)
  x_vals <- x_vals[keep]
  y_vals <- y_vals[keep]
  
  x_start <- x_vals[1];              y_start <- y_vals[1]
  x_end   <- x_vals[length(x_vals)]; y_end   <- y_vals[length(y_vals)]
  line_vec  <- c(x_end - x_start, y_end - y_start)
  norm_line <- sqrt(sum(line_vec^2))
  distances <- vapply(seq_along(x_vals), function(i) {
    point_vec <- c(x_vals[i] - x_start, y_vals[i] - y_start)
    abs(line_vec[1] * point_vec[2] - line_vec[2] * point_vec[1]) / norm_line
  }, numeric(1))
  max_index <- which.max(distances)
  elbow_x <- x_vals[max_index]
  elbow_y <- y_vals[max_index]
  
  library(ggplot2)
  library(scales)
  
  p <- ggplot(miss_cor, aes(x = missing_rate, y = mean_abundance)) +
    geom_point(size = point_size, alpha = point_alpha) +
    geom_smooth(method = "loess", se = FALSE, span = span) +
    geom_abline(
      slope = (y_end - y_start) / (x_end - x_start),
      intercept = y_start - ((y_end - y_start) / (x_end - x_start)) * x_start,
      linetype = "dashed"
    ) +
    geom_point(aes(x = elbow_x, y = elbow_y), color = "red", size = 2) +
    geom_text(
      aes(x = elbow_x, y = elbow_y + 4,
          label = paste0("Elbow @ ", scales::percent(elbow_x, accuracy = 0.001))),
      hjust = -0.2, vjust = -0.5, color = "red",
      size = 6 
    ) +
    scale_x_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, by = 0.25),
      labels = scales::percent_format(accuracy = 1)
    ) +
    scale_y_continuous(expand = expansion(mult = c(0.02, 0.05))) +
    labs(
      title = "Elbow Point in 'Dataset SCP'",
      x = "Missing Rate",
      y = "Mean log2 Abundance"
    ) +
    theme_pub(base_size = 13) +
  theme(
    axis.title = element_text(face = "bold"),   
    axis.text  = element_text(face = "bold")   
  )
  
  if (!is.null(save_path)) {
    ggsave(filename = save_path, plot = p, device = "tiff",
           dpi = tiff_dpi, width = tiff_width, height = tiff_height,
           units = "in", compression = "lzw")
  }
  
  list(
    miss_cor = miss_cor,
    fit = fit,
    elbow = list(x = elbow_x, y = elbow_y),
    plot = p
  )
}

res_scp <- find_missing_rate_elbow(df_scp, span = 0.75, tiff_width = 4.5, tiff_height = 4.5,
                                   save_path = "./figure_benchmarking/scp_elbow.tiff")

res_scp$plot
```

```{r}
soft_hybrid_impute <- function(df_rf, df_minProb, df_raw, 
                               r0 = NULL, x0 = NULL, a = 10, b = 5, lambda = 0.4,
                               visualize = TRUE) {
  stopifnot(all(dim(df_rf) == dim(df_minProb)))
  stopifnot(all(rownames(df_rf) == rownames(df_minProb)))

  missing_rate <- rowMeans(is.na(df_raw))
  mean_intensity <- rowMeans(df_raw, na.rm = TRUE)
  r0 <- elbow_x
  x0 <- elbow_y

  sigmoid <- function(x, k, x0) {
    1 / (1 + exp(-k * (x - x0)))
  }
  sigmoid_r <- sigmoid(missing_rate, k = a, x0 = r0)
  sigmoid_x <- sigmoid(mean_intensity, k = b, x0 = x0)
  p_MNAR <- sigmoid_r * (1 - lambda * sigmoid_x)

  df_imp_soft <- df_rf
  for (i in seq_len(nrow(df_rf))) {
    w <- 1 - p_MNAR[i]
    df_imp_soft[i, ] <- w * df_rf[i, ] + (1 - w) * df_minProb[i, ]
  }

  if (visualize) {
    library(ggplot2)
    df_plot <- data.frame(
      Protein = rownames(df_rf),
      MissingRate = missing_rate,
      MeanIntensity = mean_intensity,
      Weight_RF = 1 - p_MNAR
    )
     p <- ggplot(df_plot, aes(x = MissingRate, y = MeanIntensity, color = Weight_RF)) +
        geom_point(size = 0.8, alpha = 0.8) +
        scale_color_gradient(low = "blue", high = "red") +
    scale_x_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, by = 0.25),
      labels = scales::percent_format(accuracy = 1)
    ) +
    scale_y_continuous(expand = expansion(mult = c(0.02, 0.05))) +
    labs(
      title = paste("Soft Hybrid Weights in 'Dataset SCP'"),
      x = "Missing Rate",
      y = "Mean log2 Abundance"
    ) +
    theme_pub(base_size = 13) +
    theme(
      axis.title = element_text(face = "bold"),  
      axis.text  = element_text(face = "bold"),
      legend.position = c(0.85, 0.75),
      legend.title = element_text(size = 13, face = "bold"),
      legend.text  = element_text(size = 12)   
    )
    print(p)
    ggsave(
      filename = './figure_benchmarking/plot_SH.tiff', 
         plot = p, 
         device = "tiff",
         dpi = 300,
         width = 4.5, 
         height = 4.5,
         units = "in"
    )
     }
  return(df_imp_soft)
}

```

```{r}
df_imp_soft <- soft_hybrid_impute(imp_all, minProb_all, df_scp, lambda = 0.4)
save(df_imp_soft, file = './data/SoftHybridImputation_benchmarking.Rdata')
load(file = './data/SoftHybridImputation_benchmarking.Rdata')
```
## Assessment of different imputation methods
```{r}
df_truth <- df_bulk %>%
  mutate(gene = rownames(.)) %>%
  pivot_longer(cols = colnames(df_bulk), 
               names_to = "sample", 
               values_to = "intensity")
df_truth$group <- plyr::mapvalues(df_truth$sample,
                                  from = metadata_bulk$sample,
                                  to = metadata_bulk$group)
df_truth$group <- factor(df_truth$group, levels = c(1,2,3,4))
levels(df_truth$group) <- c("GSC", 'AST', "MAC", "MIC")
df_satas_truth <- df_truth %>%
  group_by(gene, group) %>%
  summarise(
    MeanExpr   = mean(intensity, na.rm = TRUE),
    .groups = "drop"
  )
NA_rm <- df_satas_truth %>%
  filter(is.na(MeanExpr)) %>%
  .$gene
NA_rm <- unique(NA_rm)
df_satas_truth <- df_satas_truth %>%
  filter(!gene %in% NA_rm)
test <- unique(df_satas_truth$gene)

imp_all_uni <- imp_all %>%
  filter(rownames(.) %in% test)
minProb_all_uni <- minProb_all %>%
  filter(rownames(.) %in% test)
knn_all_uni <- knn_all %>%
  filter(rownames(.) %in% test)
GS_all_uni <- GS_all %>%
  filter(rownames(.) %in% test)
imp_soft_uni <- df_imp_soft %>%
  filter(rownames(.) %in% test)
```
### Table Organisation
```{r}
method_levels <- c("Gaussian", "KNN", "minProb", "RF", "SoftHybrid")

method_colors <- c(
  "Gaussian"   = "#E69F00",
  "KNN"        = "#56B4E9",
  "minProb"    = "#D55E00",
  "RF"         = "#009E73",
  "SoftHybrid" = "#9966CC"
)

imp_list_uni <- list(
  Gaussian   = GS_all_uni,
  KNN        = knn_all_uni,
  minProb    = minProb_all_uni,
  RF         = imp_all_uni,
  SoftHybrid = imp_soft_uni
)

imp_list_all <- list(
  Gaussian   = GS_all,
  KNN        = knn_all,
  minProb    = minProb_all,
  RF         = imp_all,
  SoftHybrid = df_imp_soft
)

truth_mat <- df_satas_truth %>%
  pivot_wider(names_from = group, values_from = MeanExpr) %>%
  column_to_rownames("gene") %>%
  as.matrix()

metadata_use <- metadata_scp %>%
  filter(sample %in% colnames(imp_list_uni$RF)) %>%
  mutate(group = factor(group, levels = c("GSC", "AST", "MAC", "MIC")))
```
### Sample-wise AE and Spearman correlation
```{r}
evaluate_universal <- function(imputed_mat, truth_mat, metadata) {
  
  common_genes <- intersect(rownames(imputed_mat), rownames(truth_mat))
  common_samples <- intersect(colnames(imputed_mat), metadata$sample)
  
  imputed_mat <- imputed_mat[common_genes, common_samples, drop = FALSE]
  metadata_sub <- metadata %>% filter(sample %in% common_samples)
  
  result_list <- lapply(unique(metadata_sub$group), function(grp) {
    
    samples_grp <- metadata_sub %>%
      filter(group == grp) %>%
      pull(sample)
    
    truth_vec <- truth_mat[common_genes, grp]
    expr_mat  <- imputed_mat[common_genes, samples_grp, drop = FALSE]
    
    df_stat <- lapply(colnames(expr_mat), function(smp) {
      x <- expr_mat[, smp]
      data.frame(
        Sample   = smp,
        group    = grp,
        RMSE     = sqrt(mean((x - truth_vec)^2, na.rm = TRUE)),
        AE       = mean(abs(x - truth_vec), na.rm = TRUE),
        Pearson  = cor(x, truth_vec, method = "pearson",  use = "complete.obs"),
        Spearman = cor(x, truth_vec, method = "spearman", use = "complete.obs")
      )
    }) %>% bind_rows()
    
    df_stat
  })
  
  bind_rows(result_list)
}

recovery_results <- map2_dfr(imp_list_all, names(imp_list_all), ~{
  df <- evaluate_universal(.x, truth_mat, metadata_use)
  df$Method <- .y
  df
})

recovery_results$Method <- factor(recovery_results$Method, levels = method_levels)
```
### Repeated ARI and NMI
```{r}
calc_cluster_metrics_once <- function(expr_mat, metadata, k = 4, prop = 0.8, seed = 123) {
  set.seed(seed)
  
  samples_keep <- sample(metadata$sample, size = ceiling(nrow(metadata) * prop), replace = FALSE)
  meta_sub <- metadata %>% filter(sample %in% samples_keep)
  
  expr_sub <- expr_mat[, meta_sub$sample, drop = FALSE]
  
  pca <- prcomp(t(expr_sub), scale. = TRUE)
  pca_coord <- pca$x[, 1:min(10, ncol(pca$x)), drop = FALSE]
  
  km <- kmeans(pca_coord, centers = k, nstart = 50)
  
  true_label <- as.character(meta_sub$group)
  pred_label <- as.character(km$cluster)
  
  data.frame(
    ARI = adjustedRandIndex(true_label, pred_label),
    NMI = aricode::NMI(true_label, pred_label)
  )
}

calc_cluster_metrics_repeat <- function(expr_mat, metadata, k = 4, prop = 0.8, n_repeat = 20, seed = 123) {
  map_dfr(seq_len(n_repeat), function(i) {
    out <- calc_cluster_metrics_once(
      expr_mat = expr_mat,
      metadata = metadata,
      k = k,
      prop = prop,
      seed = seed + i
    )
    out$Replicate <- i
    out
  })
}

cluster_results <- map2_dfr(imp_list_all, names(imp_list_all), ~{
  df <- calc_cluster_metrics_repeat(
    expr_mat  = .x,
    metadata  = metadata_use,
    k         = 4,
    prop      = 0.8,
    n_repeat  = 100,
    seed      = 123
  )
  df$Method <- .y
  df
})

cluster_results$Method <- factor(cluster_results$Method, levels = method_levels)
```
### Long table organisation
```{r}
plot_df_recovery <- recovery_results %>%
  dplyr::select(Sample, group, Method, AE, Spearman) %>%
  pivot_longer(
    cols = c(AE, Spearman),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  mutate(
    Metric = recode(
      Metric,
      "AE" = "AE",
      "Spearman" = "Spearman"
    ),
    ID = paste(group, Sample, sep = "_")
  )

plot_df_cluster <- cluster_results %>%
  dplyr::select(Replicate, Method, ARI, NMI) %>%
  pivot_longer(
    cols = c(ARI, NMI),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  mutate(
    Metric = recode(
      Metric,
      "ARI" = "ARI",
      "NMI" = "NMI"
    ),
    ID = paste0("Rep", Replicate)
  )

plot_df_all <- bind_rows(plot_df_recovery, plot_df_cluster) %>%
  mutate(
    Method = factor(Method, levels = method_levels),
    Metric = factor(Metric, levels = c("ARI", "NMI", "AE", "Spearman"))
  )
```
### Significance Test
```{r}
compare_vs_soft <- function(df_metric) {
  
  other_methods <- setdiff(method_levels, "SoftHybrid")
  
  res <- lapply(other_methods, function(m) {
    df_wide <- df_metric %>%
      filter(Method %in% c("SoftHybrid", m)) %>%
      dplyr::select(ID, Method, Value) %>%
      distinct() %>%
      pivot_wider(names_from = Method, values_from = Value) %>%
      na.omit()
    
    if (nrow(df_wide) < 2) {
      pval <- NA_real_
      n_pair <- nrow(df_wide)
    } else {
      pval <- wilcox.test(
        df_wide[[m]],
        df_wide[["SoftHybrid"]],
        paired = TRUE,
        exact = FALSE
      )$p.value
      n_pair <- nrow(df_wide)
    }
    
    data.frame(
      group1 = m,
      group2 = "SoftHybrid",
      p = pval,
      N = n_pair
    )
  }) %>% bind_rows()
  
  res$p.adj <- p.adjust(res$p, method = "BH")
  res$p.signif <- case_when(
    is.na(res$p.adj)        ~ "ns",
    res$p.adj < 0.0001      ~ "****",
    res$p.adj < 0.001       ~ "***",
    res$p.adj < 0.01        ~ "**",
    res$p.adj < 0.05        ~ "*",
    TRUE                    ~ "ns"
  )
  
  res
}

sig_df <- plot_df_all %>%
  group_by(Metric) %>%
  group_modify(~{
    tmp_sig <- compare_vs_soft(.x)
    
    y_max <- max(.x$Value, na.rm = TRUE)
    y_min <- min(.x$Value, na.rm = TRUE)
    y_range <- y_max - y_min
    
    if (y_range == 0) y_range <- 0.1
    
    tmp_sig$y.position <- y_max + seq(0.08, by = 0.08, length.out = nrow(tmp_sig)) * y_range
    tmp_sig
  }) %>%
  ungroup()
```

### Plotting
```{r}
p_metric_box <- ggplot(
  plot_df_all,
  aes(x = Method, y = Value, fill = Method)
) +
  geom_boxplot(
    width = 0.65,
    outlier.shape = NA,
    color = "black",
    linewidth = 0.5
  ) +
  stat_summary(
    fun = mean,
    geom = "point",
    shape = 21,
    size = 2.8,
    fill = "white",
    color = "black",
    stroke = 0.8
  ) +
  stat_pvalue_manual(
    sig_df,
    label = "p.signif",
    xmin = "group1",
    xmax = "group2",
    y.position = "y.position",
    tip.length = 0.01,
    bracket.size = 0.5,
    size = 5,
    inherit.aes = FALSE
  ) +
  facet_wrap(~Metric, scales = "free_y", ncol = 4) +
  scale_fill_manual(values = method_colors) +
  labs(
    x = "Imputation Method",
    y = NULL,
    title = 'Benchmarking of SoftHybrid and commonly used methods'
  ) +
  theme_bw() +
  theme(
    panel.border      = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text              = element_text(size = 13),
    axis.text         = element_text(size = 12.5, face = "bold", color = "black"),
    axis.text.x       = element_blank(),
    axis.title        = element_text(size = 14, face = "bold"),
    axis.title.x      = element_blank(),
    axis.ticks.length.y = unit(2, "mm"),
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank(),
    legend.position   = "right",
    legend.title      = element_text(size = 13, face = "bold"),
    legend.text       = element_text(size = 12),
    plot.title        = element_text(hjust = 0.5, face = "bold", size = 15),
    strip.background  = element_rect(fill = "white", color = "black", linewidth = 0.6),
    strip.text        = element_text(size = 13, face = "bold")
  )

p_metric_box
```
```{r}
ggsave(
  filename = "./figure_benchmarking/SHI_FourMethods.tiff",  
  plot = p_metric_box,                      
  device = "tiff",                    
  dpi = 300,                         
  width = 9.5,                         
  height = 5.25,                        
  units = "in"                        
)
```


## Cell-type-specific Protein Assessment
```{r}
colnames(imp_list_all$RF) <- paste0(colnames(imp_all), '_RF')
colnames(imp_list_all$minProb) <- paste0(colnames(minProb_all), '_minProb')
colnames(imp_list_all$KNN) <- paste0(colnames(knn_all), '_KNN')
colnames(imp_list_all$Gaussian) <- paste0(colnames(GS_all), '_Gaussian')
colnames(imp_list_all$SoftHybrid) <- paste0(colnames(df_imp_soft), '_SoftHybrid')

imp_list_combined <- cbind(imp_list_all$RF,imp_list_all$minProb,imp_list_all$KNN,imp_list_all$Gaussian,imp_list_all$SoftHybrid)

metadata_combined <- data.frame(
  'Sample' = colnames(imp_list_combined),
  'Method' = c(rep(c('RF', 'minProb', 'KNN', 'Gaussian', 'SoftHybrid'), each = 74)),
  'Group' = c(rep(metadata_scp$group, 5))
)

metadata_combined$Method <- factor(metadata_combined$Method, levels = c('RF', 'minProb', 'KNN', 'Gaussian', 'SoftHybrid'))
metadata_combined$Group <- factor(metadata_combined$Group, levels = c('GSC', 'AST', 'MAC', 'MIC'))
rownames(metadata_combined) <- metadata_combined$Sample
```

```{r}
group_col <- "Group"

metadata_use <- metadata_combined %>%
  rownames_to_column("sample")

# ProteinA status
A_stats <- df_stats %>%
  filter(gene %in% marker_candidate) %>%
  mutate(
    status = case_when(
      NonNA_rate > 0.75 ~ "expressing",
      NA_rate    > 0.75 ~ "nonexpressing",
      TRUE               ~ "intermediate"
    )
  )

# bulk map for Class A proteins
A_map <- A_stats %>%
  filter(status %in% c("expressing", "nonexpressing")) %>%
  group_by(gene) %>%
  summarise(
    expressing_groups    = list(unique(group[status == "expressing"])),
    nonexpressing_groups = list(unique(group[status == "nonexpressing"])),
    bulk_expr_mean = mean(MeanExpr[status == "expressing"], na.rm = TRUE),
    bulk_expr_max  = max(MeanExpr[status == "expressing"], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(
    is.finite(bulk_expr_mean),
    !is.na(bulk_expr_mean),
    lengths(expressing_groups) > 0,
    lengths(nonexpressing_groups) > 0
  )

bulk_ref_A <- A_map %>%
  dplyr::select(
    gene,
    bulk_expr_mean,
    bulk_expr_max
  )
```

```{r}
calc_fc_imp_one_method <- function(mat, metadata_tbl, A_map_tbl,
                                   group_col = "Group",
                                   method_name = NA_character_) {
  samples_use <- intersect(colnames(mat), metadata_tbl$sample)

  if (length(samples_use) < 3) {
    warning(paste0("Method ", method_name, ": matched samples < 3; returning empty."))
    return(tibble(
      gene = character(),
      Method = character(),
      imp_expr_mean = numeric(),
      imp_nonexpr_mean = numeric(),
      FC_imp = numeric()
    ))
  }

  mat2 <- mat[, samples_use, drop = FALSE]

  md2 <- metadata_tbl[metadata_tbl$sample %in% samples_use, , drop = FALSE]
  md2 <- md2[match(samples_use, md2$sample), , drop = FALSE]

  mat2 <- mat2[, md2$sample, drop = FALSE]

  df_long <- as.data.frame(mat2) %>%
    rownames_to_column("gene") %>%
    pivot_longer(
      cols = -gene,
      names_to = "sample",
      values_to = "expr"
    ) %>%
    left_join(md2[, c("sample", group_col)], by = "sample")

  df_long$group_sc <- df_long[[group_col]]

  sc_means <- df_long %>%
    group_by(gene, group_sc) %>%
    summarise(
      mean_expr = mean(expr, na.rm = TRUE),
      .groups = "drop"
    )

  res <- A_map_tbl %>%
    dplyr::select(gene, expressing_groups, nonexpressing_groups) %>%
    rowwise() %>%
    mutate(
      imp_expr_mean = {
        eg <- unlist(expressing_groups)
        vals <- sc_means$mean_expr[
          sc_means$gene == gene & sc_means$group_sc %in% eg
        ]
        if (length(vals) == 0 || all(is.na(vals))) NA_real_ else mean(vals, na.rm = TRUE)
      },
      imp_nonexpr_mean = {
        ng <- unlist(nonexpressing_groups)
        vals <- sc_means$mean_expr[
          sc_means$gene == gene & sc_means$group_sc %in% ng
        ]
        if (length(vals) == 0 || all(is.na(vals))) NA_real_ else mean(vals, na.rm = TRUE)
      }
    ) %>%
    ungroup() %>%
    mutate(
      FC_imp = imp_expr_mean - imp_nonexpr_mean,
      Method = method_name
    ) %>%
    dplyr::select(
      gene, Method,
      imp_expr_mean, imp_nonexpr_mean, FC_imp
    )

  res
}
```

```{r}
fc_imp_all <- imap_dfr(
  imp_list_all,
  ~ calc_fc_imp_one_method(
    mat = .x,
    metadata_tbl = metadata_use,
    A_map_tbl = A_map,
    group_col = group_col,
    method_name = .y
  )
)

fc_compare_all <- fc_imp_all %>%
  left_join(bulk_ref_A, by = "gene") %>%
  mutate(
    preservation = FC_imp / bulk_expr_mean
  ) %>%
  filter(
    is.finite(FC_imp),
    is.finite(bulk_expr_mean)
  )
```

```{r}
fc_compare_all$Method <- factor(
  fc_compare_all$Method,
  levels = c("Raw", "Gaussian", "KNN", "minProb", "RF", "SoftHybrid")
)

cor_summary <- fc_compare_all %>%
  group_by(Method) %>%
  summarise(
    spearman_rho = cor(
      bulk_expr_mean, FC_imp,
      method = "spearman",
      use = "complete.obs"
    ),
    pearson_r = cor(
      bulk_expr_mean, FC_imp,
      method = "pearson",
      use = "complete.obs"
    ),
    n = sum(complete.cases(bulk_expr_mean, FC_imp)),
    .groups = "drop"
  )

cor_summary
```
```{r}
cor_label_df <- fc_compare_all %>%
  group_by(Method) %>%
  summarise(
    rho = cor(
      bulk_expr_mean, FC_imp,
      method = "spearman",
      use = "complete.obs"
    ),
    x_pos = -4,
    y_pos = 5,
    .groups = "drop"
  ) %>%
  mutate(
    label = paste0("rho = ", round(rho, 3))
  )

method_colors <- c(
  "Gaussian"   = "#E69F00",
  "KNN"        = "#56B4E9",
  "minProb"    = "#D55E00",
  "RF"         = "#009E73",
  "SoftHybrid" = "#9966CC"
)

p_fc_vs_bulkexpr_lab <- ggplot(
  fc_compare_all,
  aes(x = bulk_expr_mean, y = FC_imp, color = Method)
) +
  geom_point(alpha = 0.7, size = 1.5) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
  geom_text(
    data = cor_label_df,
    aes(x = x_pos, y = y_pos, label = label),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1,
    size = 5.5,
    fontface = "bold"
  ) +
  scale_color_manual(values = method_colors) +
  facet_wrap(~ Method, scales = "fixed", ncol = 3) +
  labs(
    x = "Bulk mean expression in expressing cell type(s)",
    y = "Imputed FC 
    (expressing vs non-expressing)"
  ) +
  theme_classic(base_size = 12) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
  

p_fc_vs_bulkexpr_lab

ggsave(
  filename = "./figure_benchmarking/FC_abundance_scatter.tiff",  
  plot = p_fc_vs_bulkexpr_lab,                      
  device = "tiff",                   
  dpi = 300,                         
  width = 9,                        
  height = 5.5,                       
  units = "in"                     
)
```
# Four Imputation Method: Current Gap
## PCA plot
```{r}
imp_list <- list(
  RF     = imp_all,
  minProb = minProb_all,
  KNN    = knn_all,
  Gaussian  = GS_all
)

colnames(imp_list$RF) <- paste0(colnames(imp_all), '_RF')
colnames(imp_list$minProb) <- paste0(colnames(minProb_all), '_minProb')
colnames(imp_list$KNN) <- paste0(colnames(knn_all), '_KNN')
colnames(imp_list$Gaussian) <- paste0(colnames(GS_all), '_Gaussian')

imp_list_combined <- cbind(imp_list$RF,imp_list$minProb,imp_list$KNN,imp_list$Gaussian)

metadata_combined <- data.frame(
  'Sample' = colnames(imp_list_combined),
  'Method' = c(rep(c('RF', 'minProb', 'KNN', 'Gaussian'), each = 74)),
  'Group' = c(rep(metadata_scp$group, 4))
)

metadata_combined$Method <- factor(metadata_combined$Method, levels = c('RF', 'minProb', 'KNN', 'Gaussian'))
metadata_combined$Group <- factor(metadata_combined$Group, levels = c('GSC', 'AST', 'MAC', 'MIC'))
rownames(metadata_combined) <- metadata_combined$Sample
```

```{r,warning=FALSE}
pca <- prcomp(x = t(imp_list_combined), center = T)
std_explained <- pca$sdev^2 / sum(pca$sdev^2)

PC <- as.data.frame(pca$x) 
PC$sample <- rownames(PC)
PC <- merge(PC, metadata_combined, by.x = "sample", by.y = "row.names") %>%
  dplyr::select(-sample)

xlab <- paste0("PC1: ",round(std_explained[1]*100,1),'%')
ylab <- paste0("PC2: ",round(std_explained[2]*100,1),'%')

n_proteins <- nrow(imp_list_combined)
title <- paste0("Comparison of three cell types")
my_colors <- c(
  "RF"         = "#009E73",  
  "minProb"    = "#D55E00",  
  "Gaussian"   = "#E69F00",  
  "KNN"        = "#56B4E9",  
  "SoftHybrid" = "#9966CC"  )

p1 <- ggplot(data = PC, aes(x=PC1, y=PC2, color=Method, shape = Group)) +
  geom_point(size = 2) +
  scale_color_manual(values = my_colors) + 
  theme_bw() +
  xlab(xlab) +
  ylab(ylab) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
p1

ggsave(
  filename = "./figure_benchmarking/PCA_FourMethods.tiff",  
  plot = p1,                      
  device = "tiff",                    
  dpi = 300,                        
  width = 5,                         
  height = 4,                       
  units = "in"                        
)
```
## PC loadings
```{r}
loadings <- pca$rotation[,1]
# raw missing rate
miss_rate <- rowMeans(is.na(df_scp))

df_loading <- data.frame(
  protein = names(loadings),
  loading = loadings,
  abs_loading = abs(loadings),
  miss_rate = miss_rate[names(loadings)]
)

ggplot(df_loading, aes(x = miss_rate, y = abs_loading)) +
  geom_point(alpha=0.4) +
  geom_smooth(method="lm", col="red") +
  labs(x="Missing rate (raw data)",
       y="Absolute PC1 loading")
```
```{r}
# Nature-style: PC1 loading vs missing rate (clean + density contour + stats)

# df_loading must contain:
#   miss_rate   : missing rate in raw data (0-1)
#   abs_loading : absolute PC1 loading
stopifnot(all(c("miss_rate", "abs_loading") %in% colnames(df_loading)))

# stats
r_val <- cor(df_loading$abs_loading, df_loading$miss_rate, use = "complete.obs", method = "pearson")
p_val <- cor.test(df_loading$abs_loading, df_loading$miss_rate, method = "pearson")$p.value
lab   <- paste0("Pearson r = ", sprintf("%.3f", r_val), "\n",
                "P ", ifelse(p_val < 1e-300, "< 1e-300", paste0("= ", format.pval(p_val, digits = 2, eps = 1e-300))))

# dynamic label position (top-left, avoids trend line)
x_lab <- quantile(df_loading$miss_rate, 0.03, na.rm = TRUE)
y_lab <- quantile(df_loading$abs_loading, 0.98, na.rm = TRUE)

p <- ggplot(df_loading, aes(x = miss_rate, y = abs_loading)) +
  # dense point cloud (light)
  geom_point(alpha = 0.12, size = 0.8, color = "black",
             position = position_jitter(width = 0.002, height = 0)) +
  # 2D density contours (Nature-ish)
  #geom_density_2d(linewidth = 0.35, color = "black", alpha = 0.55) +
  # linear fit (minimal, not overbearing)
  geom_smooth(method = "lm", se = TRUE,
              linewidth = 0.7, color = "#D7263D", fill = "#F4A7A7") +
  # stats label
  annotate("label",
           x = x_lab, y = y_lab, label = lab,
           hjust = 0, vjust = 1,
           label.size = 0.25, fill = "white",
           size = 4.2, family = "sans") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25), expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  labs(
    title = "PC1 driven by proteins with high MR",
    x = "Missing rate (raw data)",
    y = "Absolute PC1 loading"
  ) +
  theme_classic(base_size = 14) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )

print(p)
```

## Mantel test
```{r}
library(vegan)
library(dplyr)
library(ggplot2)
library(forcats)

# 1) raw observed distance: use pairwise complete correlation (cell-cell)
cor_raw <- cor(df_scp, use = "pairwise.complete.obs", method = "pearson")
D_raw   <- as.dist(1 - cor_raw)

# 2) Mantel per method
mantel_df <- lapply(names(imp_list), function(m){
  cor_m <- cor(imp_list[[m]], use = "everything", method = "pearson")  # no NA expected
  D_m   <- as.dist(1 - cor_m)

  mt <- vegan::mantel(D_raw, D_m, method = "spearman", permutations = 999)

  data.frame(
    method   = m,
    mantel_r = unname(mt$statistic),
    p        = mt$signif
  )
}) %>% bind_rows() %>%
  mutate(
    method = fct_reorder(method, mantel_r),
    p_lab  = "P<0.001"
  )

# 3) Visualization: ranked bar + point + p label
p_mantel <- 
ggplot(mantel_df, aes(x = method, y = mantel_r, fill = method)) +
  geom_col(width = 0.72, alpha = 0.85) +
  geom_point(size = 2.2) +
  geom_text(aes(label = p_lab), hjust = -0.05, size = 3.8) +
  coord_flip() +
  scale_fill_manual(values = my_colors) +
  scale_y_continuous(
    limits = c(min(0, min(mantel_df$mantel_r) - 0.05),
               max(mantel_df$mantel_r) + 0.08)
  ) +
  labs(
    x = NULL,
    y = "Mantel correlation (Spearman r)",
    title = "Imputation methods preserve raw cell–cell geometry"
  ) +
  theme_classic(base_size = 14) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
p_mantel
```

## Separated PCA plot
```{r,warning=FALSE}
my_colors <- c(
  "RF"         = "#009E73",
  "minProb"    = "#D55E00",
  "Gaussian"   = "#E69F00",
  "KNN"        = "#56B4E9",
  "SoftHybrid" = "#9966CC"
)

# PCA Plotting Function
plot_pca_one_method <- function(method_name,
                                imp_list,
                                metadata_df,
                                center = TRUE,
                                scale. = FALSE) {

  mat <- imp_list[[method_name]]

  samples_use <- intersect(colnames(mat), rownames(metadata_df))
  mat <- mat[, samples_use, drop = FALSE]

  sample_order <- rownames(metadata_df)[rownames(metadata_df) %in% samples_use]
  mat <- mat[, sample_order, drop = FALSE]

  pca <- prcomp(t(mat), center = center, scale. = scale.)

  std_explained <- pca$sdev^2 / sum(pca$sdev^2)

  PC <- as.data.frame(pca$x)
  PC$sample <- rownames(PC)

  PC <- merge(PC, metadata_df, by.x = "sample", by.y = "row.names") %>%
    dplyr::select(-sample)

  xlab <- paste0("PC1: ", round(std_explained[1] * 100, 1), "%")
  ylab <- paste0("PC2: ", round(std_explained[2] * 100, 1), "%")

  ggplot(PC, aes(x = PC1, y = PC2, shape = Group)) +
    stat_ellipse(
      aes(group = Group),
      type = "t",
      linewidth = 0.8,
      linetype = 1,
      color = my_colors[[method_name]],
      level = 0.95
    ) +
    geom_point(
      size = 2.5,
      color = my_colors[[method_name]]
    ) +
    theme_bw() +
    xlab(xlab) +
    ylab(ylab) +
    ggtitle(paste0("PCA — ", method_name)) +
    theme(
      panel.border      = element_rect(colour = "black", fill = NA, linewidth = 0.6),
      text              = element_text(size = 13),
      axis.text         = element_text(size = 13, face = "bold"),
      axis.text.x       = element_text(color = "black", size = 13, face = "bold"),
      axis.text.y       = element_text(color = "black", size = 13, face = "bold"),
      axis.ticks.length = unit(2, "mm"),
      axis.title        = element_text(size = 14, face = "bold"),
      panel.grid.major  = element_blank(),
      panel.grid.minor  = element_blank(),
      legend.position   = "right",
      legend.title      = element_text(size = 13, face = "bold"),
      legend.text       = element_text(size = 12),
      plot.title        = element_text(hjust = 0.5, face = "bold", size = 15)
    )
}

methods_use <- names(imp_list)

p_list <- lapply(
  methods_use,
  plot_pca_one_method,
  imp_list = imp_list,
  metadata_df = metadata_combined
)

names(p_list) <- methods_use

(p_list[["RF"]] | p_list[["minProb"]]) /
  (p_list[["Gaussian"]] | p_list[["KNN"]])

methods_use <- names(imp_list)

p_list <- lapply(methods_use,
                 plot_pca_one_method,
                 imp_list = imp_list,
                 metadata_df = metadata_combined)

names(p_list) <- methods_use

combo_pca <- (p_list[["RF"]] | p_list[["minProb"]]) / (p_list[["Gaussian"]] | p_list[["KNN"]])
combo_pca

combo <- (p / p_mantel)| combo_pca+
  plot_annotation(tag_levels = "A") &
  theme(
    plot.tag = element_text(face = "bold", size = 16)
  )
combo

ggsave(
  filename = "./figure_benchmarking/reviewer1_1.tiff",  
  plot = combo,                      
  device = "tiff",                    
  dpi = 300,                         
  width = 15,                         
  height = 5.25,                        
  units = "in"                        
)
```

## ARI/Cor Comparison
```{r}
calc_ari <- function(expr_mat, metadata, k = 4) {
  pca <- prcomp(t(expr_mat), scale. = TRUE)
  pca_coord <- pca$x  
  
  set.seed(123)
  km <- kmeans(pca_coord, centers = k, nstart = 50)
  
  true_label <- metadata$group
  pred_label <- km$cluster
  ari <- adjustedRandIndex(true_label, pred_label)
  
  return(ari)
}

methods_lv <- c("RF", "minProb", "KNN", "Gaussian")
types_lv   <- c("GSC", "AST", "MAC", "MIC")

meta <- metadata_combined %>%
  mutate(
    method = factor(Method, levels = methods_lv),
    group  = factor(Group,  levels = types_lv)
  )

compute_within_cor_for_method <- function(expr_mat, labels) {
  C <- suppressWarnings(cor(expr_mat, method = "pearson", use = "pairwise.complete.obs"))
  smp <- colnames(expr_mat)
  lab <- labels[smp]

  n <- length(smp)
  idx <- which(upper.tri(matrix(NA, n, n)), arr.ind = TRUE)
  same_grp <- lab[idx[,1]] == lab[idx[,2]]
  vals <- C[cbind(idx[same_grp,1], idx[same_grp,2])]

  if (length(vals) == 0) return(NA_real_)
  stats::median(vals, na.rm = TRUE)
}

res_list <- lapply(methods_lv, function(m) {
  smps_m <- meta %>% filter(method == m) %>% pull(Sample)
  
  if (!all(smps_m %in% colnames(imp_list_combined))) {
    stop(sprintf("Some samples for method %s are not found in imp_list_combined.", m))
  }
  
  expr_m <- imp_list_combined[, smps_m, drop = FALSE]
  
  meta_m <- meta %>% filter(Method == m)
  
  labs_m <- meta_m %>% pull(Group) %>% as.character()
  names(labs_m) <- meta_m$Sample
  
  ari_m <- calc_ari(expr_m, meta_m, k = 4)
  cor_m <- compute_within_cor_for_method(expr_m, labs_m)
  
  tibble(Method = m, ARI = ari_m, Cor = cor_m)
})

res <- bind_rows(res_list)

res_long <- res %>%
  pivot_longer(cols = c(ARI, Cor), names_to = "Metric", values_to = "Value") %>%
  mutate(
    Method = factor(Method, levels = methods_lv),
    Metric = factor(Metric, levels = c("ARI", "Cor"))
  )

print(res)
print(res_long)
```

```{r}
res_ARI <- res$ARI
res_cor <- res$Cor

zscore <- function(x) {
  if (sd(x) == 0) {
    return(rep(0, length(x)))  # 避免标准差为0
  } else {
    (x - mean(x)) / sd(x)
  }
}

res_ARI_z <- zscore(res_ARI)
res_cor_z <- zscore(res_cor)

df_z <- data.frame(
  Method = res$Method,
  ARI_z = res_ARI_z,
  Cor_z = res_cor_z
)

df_long_z <- df_z %>%
  tidyr::pivot_longer(cols = c(ARI_z, Cor_z),
                      names_to = "Metric", values_to = "Zscore")

library(ggplot2)
pz <- 
  ggplot(df_long_z, aes(x = Method, y = Zscore, fill = Metric)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  scale_fill_manual(values = c("ARI_z" = "#7B9ACC", "Cor_z" = "#F2A65A"),
                    labels = c("ARI", "RepCorrelation")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  
  theme_bw() +
  labs(title = "Z-scored ARI and Cor by method",
       x = "Imputation method", y = "Z-score", fill = NULL) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"),
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pz

ggsave(
  filename = "./figure_benchmarking/ARI_Cor_FourMethods.tiff",  
  plot = pz,                      
  device = "tiff",                   
  dpi = 300,                         
  width = 7,                        
  height = 4,                       
  units = "in"                     
)
```

```{r}
sessionInfo()
```

