---
title: "4_ImputationBenchmarking_SCP_Condition_DIA"
author: "Yixin Shi"
date: "2025-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(magrittr, ggplot2, ggVennDiagram, pheatmap, ggridges, umap, limma, dplyr, tidyr, viridis, gridExtra, stringr, ggsci, clusterProfiler, org.Mm.eg.db, UpSetR, missForest,ComplexHeatmap, WGCNA, org.Hs.eg.db, GOSemSim, ggrepel, VennDiagram, GSVA, tibble, purrr, scales, imputeLCMD, patchwork)
```

# Data Preperation
## Read Data
```{r, warning=FALSE}
rm(list = ls())
getwd()

intensity <- read.csv("./data/report.unique_genes_matrix_con_rerun.tsv", sep = "\t", header = T)

rownames(intensity) <- intensity$Genes
intensity <- intensity[,-c(1,2,3)]
sample <- unlist(lapply(strsplit(colnames(intensity), split = '_', fixed = TRUE), function(x) x[12]))
sample <- unlist(lapply(strsplit(sample, split = '.', fixed = TRUE), function(x) x[1]))
group <- unlist(lapply(strsplit(colnames(intensity), split = '.', fixed = TRUE), function(x) x[7]))
group <- unlist(lapply(strsplit(group, split = '_', fixed = TRUE), function(x) paste0(x[1], '_', x[2])))

metadata <- data.frame(
  Sample = sample,
  Group = group
)
metadata$Sample[1:60] <- paste0(metadata$Sample[1:60],'_H')
metadata$Sample[61:120] <- paste0(metadata$Sample[61:120],'_N')
rownames(metadata) <- metadata$Sample
colnames(intensity) <- metadata$Sample

intensity <- cbind(intensity[,81:100], intensity[,c(61:80, 101:120)], intensity[, 21:40], intensity[,c(1:20,41:60)])
metadata <- metadata[colnames(intensity),]
```

## Prepare Metadata File
```{r, warning=FALSE}
metadata$Condition <- unlist(lapply(strsplit(metadata$Group, split = '_', fixed = TRUE), function(x) x[1]))
metadata$Type <- unlist(lapply(strsplit(metadata$Group, split = '_', fixed = TRUE), function(x) x[2]))
metadata$Condition <- factor(metadata$Condition, levels = c('Normoxia', 'Hypoxia'))
metadata$Type <- factor(metadata$Type, levels = c('GSC', 'AST', 'MAC'))
metadata$ID <- apply(intensity, 2, function(x) sum(!is.na(x)))
```

## Data matrix manipulation
```{r}
df_long <- intensity %>% 
  mutate(gene = rownames(.)) %>%
  pivot_longer(cols = colnames(intensity), 
               names_to = "sample", 
               values_to = "intensity")
df_long$group <- plyr::mapvalues(df_long$sample,
                                       from = metadata$Sample,
                                       to = metadata$Group)
df_long$group <- factor(df_long$group, levels = c('Hypoxia_GSC', 'Hypoxia_AST', 'Hypoxia_MAC', 'Normoxia_GSC', 'Normoxia_AST', 'Normoxia_MAC'))
df_long$sample <- factor(df_long$sample, levels = metadata$Sample)
```

## Filter expression profile in merged dataset
```{r}
### Manipulate data by log2
df_log2 <- log2(intensity)
df_long_log2 <- df_long
df_long_log2$intensity <- log2(df_long_log2$intensity)
### Regroup the matrix based on group and gene
condition <- df_long %>%  
  group_by(gene, group)%>%
  reframe(sum(!is.na(intensity))) %>%
  ungroup()
colnames(condition) <- c('gene', 'group', 'nonna')

metadata$Sample <- as.character(metadata$Sample)
filter_sample <- metadata[metadata$ID >= 1000, "Sample"]
metadata <- metadata[filter_sample,]
```

```{r}
### Filter those gene which was detected in more than 75% replicates in at least one group
geneName1 <- vector()
for (i in unique(condition$gene)) {
  if(any(condition[condition$gene == i,]$nonna>15)) {
    geneName1 <- c(geneName1, i)
    }
}
df_long_filter <- df_long_log2 %>%
  filter(gene %in% geneName1) %>%
  filter(sample %in% filter_sample)

df_filter <- df_long_filter [,-4] %>%
  pivot_wider(names_from = sample,
              values_from = intensity) %>%
  as.data.frame()
rownames(df_filter )<- df_filter$gene
df_filter <- df_filter[,-1]
```

##  Normlisation
```{r}
### Data normalization function
median_centering <- function(df) {
  df_norm <- as.data.frame(apply(df, 2, function(x) {
                        gMedian = median(x, na.rm = TRUE)
                        x - gMedian
  }
    ))
  df_long_norm <- df_norm %>%
    mutate(gene = rownames(.)) %>%
    pivot_longer(cols = colnames(df_norm),
                 names_to = 'sample',
                 values_to = 'intensity')
  df_long_norm$group <- plyr::mapvalues(df_long_norm$sample,
                                        from = metadata$Sample,
                                        to = metadata$Group)
  df_long_norm$sample <- factor(df_long_norm$sample, levels = metadata$Sample)
  
  return(df_long_norm)
}
### Normalize data
df_long_norm <- median_centering(df_filter)
df_long_norm$group <- factor(df_long_norm$group, levels = c('Hypoxia_GSC', 'Hypoxia_AST', 'Hypoxia_MAC', 'Normoxia_GSC', 'Normoxia_AST', 'Normoxia_MAC'))

df_norm <- df_long_norm %>% 
  dplyr::select(-group) %>%
  pivot_wider(names_from = sample, values_from = intensity) %>% 
  as.data.frame()
rownames(df_norm) <- df_norm$gene
df_norm <- df_norm[,-1]
```

# Imputation
## Imputation Method I: RF
```{r}
#set.seed(123)
#res_all  <- missForest(as.data.frame(df_norm), verbose=FALSE)
#imp_all  <- res_all$ximp
#save(imp_all, file = './data_condition/imp_all_benchmarking.Rdata')
load(file = './data_condition/imp_all_benchmarking.Rdata')
```

## Imputation Method II: Minimal Probability of ALL proteins
```{r}
#set.seed(25)
#minProb_all <- impute.MinProb(
#  df_norm,
#  q = 0.01,
#  tune.sigma = 1
#)
#save(minProb_all, file = './data_condition/minProb_all_benchmarking.Rdata')
load(file = './data_condition/minProb_all_benchmarking.Rdata')
```

## Imputation Method III: KNN of ALL proteins
```{r}
#set.seed(25)
#knn_impute <- impute.knn(as.matrix(df_norm), k = 5)
#knn_all <- as.data.frame(knn_impute[[1]])
#save(knn_all, file = './data_condition/knn_all_benchmarking.Rdata')
load(file = './data_condition/knn_all_benchmarking.Rdata')
```

## Imputation Method IV: gaussian sampling impute of ALL proteins
```{r}
#set.seed(25)

#gaussian_impute <- function(df, width = 0.3, downshift = 1.8) {
#  imputed_df <- df
#  all_values <- unlist(df)
#  observed_values <- all_values[!is.na(all_values)]
#  mu <- mean(observed_values, na.rm = TRUE)
#  sigma <- sd(observed_values, na.rm = TRUE)
#  impute_mean <- mu - downshift * sigma
#  impute_sd <- sigma * width
#  for (i in 1:nrow(df)) {
#    for (j in 1:ncol(df)) {
#      if (is.na(df[i, j])) {
#        imputed_df[i, j] <- rnorm(1, mean = impute_mean, sd = impute_sd)
#      }
#    }
#  }
#  return(imputed_df)
#}

#GS_all <- gaussian_impute(df_norm)

#save(GS_all, file = './data_condition/GS_all_benchmarking.Rdata')
load(file = './data_condition/GS_all_benchmarking.Rdata')
```

## Imputation Method V: Soft Hybrid Imputation
### Assessment of missing rate
```{r}
miss_cor <- data.frame(
  missing_rate    = rowSums(is.na(df_norm)) / ncol(df_norm),
  mean_abundance  = rowMeans(df_filter, na.rm = TRUE)
)
ggplot(miss_cor, aes(x = missing_rate, y = mean_abundance)) +
  geom_point(size = 0.7, alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  theme_bw() +
  xlab("Missing Rate") +
  ylab("Mean log2 Abundance") +
  ggtitle("Correlation of Abundance and Missing Rate")
```

```{r}
loess_fit <- loess(mean_abundance ~ missing_rate, data = miss_cor, span = 0.75)

x_vals <- seq(min(miss_cor$missing_rate), max(miss_cor$missing_rate), length.out = 1000)
y_vals <- predict(loess_fit, newdata = data.frame(missing_rate = x_vals))

x_start <- x_vals[1]; y_start <- y_vals[1]
x_end   <- x_vals[length(x_vals)]; y_end <- y_vals[length(y_vals)]
line_vec <- c(x_end - x_start, y_end - y_start)

distances <- sapply(1:length(x_vals), function(i) {
  point_vec <- c(x_vals[i] - x_start, y_vals[i] - y_start)
  cross_prod <- abs(line_vec[1] * point_vec[2] - line_vec[2] * point_vec[1])
  norm_line <- sqrt(sum(line_vec^2))
  return(cross_prod / norm_line)
})

max_index <- which.max(distances)
elbow_x <- x_vals[max_index]
elbow_y <- y_vals[max_index]

ggplot(miss_cor, aes(x = missing_rate, y = mean_abundance)) +
  geom_point(size = 0.7, alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.75, color = "blue") +
  geom_abline(
    slope = (y_end - y_start) / (x_end - x_start),
    intercept = y_start - ((y_end - y_start) / (x_end - x_start)) * x_start,
    color = "gray50", linetype = "dashed"
  ) +
  geom_point(aes(x = elbow_x, y = elbow_y), color = "red", size = 2) +
  geom_text(aes(x = elbow_x, y = elbow_y,
                label = paste0("Elbow @ ", percent(elbow_x, accuracy = 1))),
            hjust = -0.2, vjust = -0.5, color = "red") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  theme_bw() +
  xlab("Missing Rate") +
  ylab("Mean log2 Abundance") +
  ggtitle("Loess-Based Elbow Point by Maximum Distance Method")
```
```{r}
soft_hybrid_impute <- function(df_rf, df_minProb, df_raw, 
                               r0 = NULL, x0 = NULL, a = 10, b = 5, lambda = 0.4,
                               visualize = TRUE) {
  stopifnot(all(dim(df_rf) == dim(df_minProb)))
  stopifnot(all(rownames(df_rf) == rownames(df_minProb)))

  missing_rate <- rowMeans(is.na(df_raw))
  mean_intensity <- rowMeans(df_raw, na.rm = TRUE)
  r0 <- elbow_x
  x0 <- elbow_y

  sigmoid <- function(x, k, x0) {
    1 / (1 + exp(-k * (x - x0)))
  }
  sigmoid_r <- sigmoid(missing_rate, k = a, x0 = r0)
  sigmoid_x <- sigmoid(mean_intensity, k = b, x0 = x0)
  p_MNAR <- sigmoid_r * (1 - lambda * sigmoid_x)

  df_imp_soft <- df_rf
  for (i in seq_len(nrow(df_rf))) {
    w <- 1 - p_MNAR[i]  
    df_imp_soft[i, ] <- w * df_rf[i, ] + (1 - w) * df_minProb[i, ]
  }

  if (visualize) {
    library(ggplot2)
    df_plot <- data.frame(
      Protein = rownames(df_rf),
      MissingRate = missing_rate,
      MeanIntensity = mean_intensity,
      Weight_RF = 1 - p_MNAR
    )
    theme_pub <- function(base_size = 10, base_family = "Arial") {
              ggplot2::theme_classic(base_size = base_size, base_family = base_family) %+replace%
                ggplot2::theme(
                  panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
                  axis.title       = element_text(size = base_size + 1, face = "bold"),
                  axis.text        = element_text(size = base_size),
                  axis.ticks.length= unit(2, "mm"),
                  plot.title       = element_text(size = base_size + 2, face = "bold", hjust = 0.5),
                  legend.position  = "right",
                  legend.title     = element_text(size = base_size),
                  legend.text      = element_text(size = base_size - 1)
                )
            }
     p <- ggplot(df_plot, aes(x = MissingRate, y = MeanIntensity, color = Weight_RF)) +
        geom_point(size = 0.8, alpha = 0.8) +
        scale_color_gradient(low = "blue", high = "red") +
    scale_x_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, by = 0.25),
      labels = scales::percent_format(accuracy = 1)
    ) +
    scale_y_continuous(expand = expansion(mult = c(0.02, 0.05))) +
    labs(
      title = paste("Soft Hybrid Weights"),
      x = "Missing Rate",
      y = "Mean log2 Abundance"
    ) +
    theme_pub(base_size = 10, base_family = "Arial")
    print(p)
    ggsave(
      filename = './figure_benchmarking/plot_SH_Con.tiff', 
         plot = p, 
         device = "tiff",
         dpi = 300,
         width = 6, 
         height = 3.5,
         units = "in"
    )
     }
  return(df_imp_soft)
}
```

```{r}
df_imp_soft <- soft_hybrid_impute(imp_all, minProb_all, df_filter, lambda = 0.4)
save(df_imp_soft, file = './data_condition/SoftHybridImputation_benchmarking.Rdata')
load(file = './data_condition/SoftHybridImputation_benchmarking.Rdata')
```

# Limma Comaprison
## Run in the same batch
```{r}
run_limma <- function(imputed_data, metadata, method_name, fc_threshold = 1.5, pval_threshold = 0.05) {
  design <- model.matrix(~0 + Condition, metadata)
  colnames(design) <- c("Normoxia", "Hypoxia")

  fit1 <- lmFit(imputed_data, design = design, method = "robust", maxit = 2000)
  contrast <- makeContrasts(Hypoxia - Normoxia, levels = colnames(design))
  fit2 <- contrasts.fit(fit1, contrast)
  fit3 <- eBayes(fit2)

  modtest <- limma::topTable(fit3, number = Inf, sort.by = "none")
  limma_result <- merge(as.data.frame(imputed_data), as.data.frame(modtest), by = "row.names", all = TRUE)
  rownames(limma_result) <- limma_result$Row.names
  limma_result <- limma_result %>%
    dplyr::select(-Row.names) %>%
    arrange(adj.P.Val) %>%
    mutate(gene = rownames(.))
  sig_res <- limma_result %>%
    filter(adj.P.Val < pval_threshold, AveExpr > 0) %>%
    arrange(adj.P.Val)
  mapID <- as.data.frame(mapIds(org.Hs.eg.db, sig_res$gene, 'ENTREZID', 'SYMBOL'))
  colnames(mapID) <- c('ENTREZID')
  mapID <- mapID %>% mutate(SYMBOL = rownames(.))
  sig_res$ENTREZID <- mapID$ENTREZID
  fold_change <- sig_res$logFC
  names(fold_change) <- sig_res$gene

  top_markers_FC_up <- sort(fold_change, decreasing = TRUE)
  top_markers_FC_down <- sort(fold_change, decreasing = FALSE)

  pvalue <- sig_res$adj.P.Val
  names(pvalue) <- sig_res$gene
  top_markers_p <- sort(pvalue, decreasing = FALSE)
  limma_result_upregulated <- sig_res %>% filter(logFC > fc_threshold)

  return(list(
    method = method_name,
    full_result = limma_result,
    significant_DEGs = sig_res,
    top_up = top_markers_FC_up,
    top_down = top_markers_FC_down,
    top_p = top_markers_p,
    upregulated = limma_result_upregulated
  ))
}
```

```{r}
results_list <- list(
  RF = run_limma(imp_all, metadata, "RF"),
  minProb = run_limma(minProb_all, metadata, "minProb"),
  SoftHybrid = run_limma(df_imp_soft, metadata, "SoftHybrid"),
  KNN = run_limma(knn_all, metadata, "KNN"),
  Gaussian = run_limma(GS_all, metadata, "Gaussian")
)
```

```{r}
hallmarks <- read.gmt('./resource/h.all.v2023.2.Hs.symbols.gmt')

run_gsea_hallmark <- function(limma_df, method_name = "") {
  gene_df <- limma_df %>%
    dplyr::select(gene, logFC) %>%
    filter(!is.na(gene), !is.na(logFC))
  geneList <- gene_df$logFC
  names(geneList) <- gene_df$gene
  geneList <- sort(geneList, decreasing = TRUE)

  if (length(geneList) < 10) {
    warning(paste("Too few genes for GSEA:", method_name))
    return(NULL)
  }

  gsea_res <- tryCatch({
    GSEA(geneList, TERM2GENE = hallmarks, pvalueCutoff = 0.25)
  }, error = function(e) NULL)

  return(gsea_res)
}
```

```{r, warning=FALSE}
set.seed(123)
gsea_results <- lapply(names(results_list), function(m) {
  run_gsea_hallmark(results_list[[m]]$full_result, method_name = m)
})
names(gsea_results) <- names(results_list)
```

```{r}
pRF <- dotplot(gsea_results$RF,
        showCategory = 20,
        label_format = 60,
        split = ".sign") +
  facet_grid(~.sign) +
  ggtitle("GSEA - RF") +
  coord_cartesian(xlim = c(0, 0.65), clip = "off") +
  scale_color_gradient2(
    high = "#4575b4", mid = "white", low = "#d73027", midpoint = 0.1,
    name = "p.adjust",
    guide = guide_colorbar(order = 1)
  ) +
  scale_size_continuous(
    name = "Count",
    guide = guide_legend(order = 2)
  ) +
  guides(
      colour = guide_colorbar(order = 1),
      size   = guide_legend(order = 2)
    ) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pRF

pMP <- dotplot(gsea_results$minProb,
        showCategory = 20,
        label_format = 60,
        split = ".sign") +
  facet_grid(~.sign) +
  ggtitle("GSEA - minProb") +
  coord_cartesian(xlim = c(0, 0.65), clip = "off") +
  scale_color_gradient2(
    high = "#4575b4", mid = "white", low = "#d73027", midpoint = 0.1,
    name = "p.adjust",
    guide = guide_colorbar(order = 1)
  ) +
  scale_size_continuous(
    name = "Count",
    guide = guide_legend(order = 2)
  ) +
  guides(
      colour = guide_colorbar(order = 1),
      size   = guide_legend(order = 2)
    ) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pMP

pSH <- dotplot(gsea_results$SoftHybrid,
        showCategory = 20,
        label_format = 60,
        split = ".sign") +
  facet_grid(~.sign) +
  ggtitle("GSEA - SoftHybrid") +
  coord_cartesian(xlim = c(0, 0.65), clip = "off") +
  scale_color_gradient2(
    high = "#4575b4", mid = "white", low = "#d73027", midpoint = 0.1,
    name = "p.adjust",
    guide = guide_colorbar(order = 1)
  ) +
  scale_size_continuous(
    name = "Count",
    guide = guide_legend(order = 2)
  ) +
  guides(
      colour = guide_colorbar(order = 1),
      size   = guide_legend(order = 2)
    ) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pSH

pKNN <- dotplot(gsea_results$KNN,
        showCategory = 20,
        label_format = 60,
        split = ".sign") +
  facet_grid(~.sign) +
  ggtitle("GSEA - KNN") +
  coord_cartesian(xlim = c(0, 0.65), clip = "off") +
  scale_color_gradient2(
    high = "#4575b4", mid = "white", low = "#d73027", midpoint = 0.1,
    name = "p.adjust",
    guide = guide_colorbar(order = 1)
  ) +
  scale_size_continuous(
    name = "Count",
    guide = guide_legend(order = 2)
  ) +
  guides(
      colour = guide_colorbar(order = 1),
      size   = guide_legend(order = 2)
    ) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pKNN

pG <- dotplot(gsea_results$Gaussian,
        showCategory = 20,
        label_format = 60,
        split = ".sign") +
  facet_grid(~.sign) +
  ggtitle("GSEA - Gaussian") +
  coord_cartesian(xlim = c(0, 0.65), clip = "off") +
  scale_color_gradient2(
    high = "#4575b4", mid = "white", low = "#d73027", midpoint = 0.1,
    name = "p.adjust",
    guide = guide_colorbar(order = 1)
  ) +
  scale_size_continuous(
    name = "Count",
    guide = guide_legend(order = 2)
  ) +
  guides(
      colour = guide_colorbar(order = 1),
      size   = guide_legend(order = 2)
    ) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pG
```
```{r}
pDot_Combo <- (pRF|pKNN)/(pMP|pG) +   
  #plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') &   
  theme(
    plot.tag = element_text(face = "bold", size = 14),
    plot.tag.position = c(0.02, 0.98)         
  )
pDot_Combo

ggsave(
  filename = "./figure_benchmarking/SCP_Condition_GSEA_DotSupp.tiff",  
  plot = pDot_Combo,                      
  device = "tiff",                    
  dpi = 300,                         
  width = 18,                         
  height = 9,                        
  units = "in"                       
)

ggsave(
  filename = "./figure_benchmarking/SCP_Condition_GSEA_DotSH.tiff",  
  plot = pSH,                      
  device = "tiff",                    
  dpi = 300,                          
  width = 9,                        
  height = 4,                       
  units = "in"                       
)
```


```{r}
metrics <- c("enrichmentScore", "NES", "pvalue", "p.adjust")

hypoxia_summary <- lapply(names(gsea_results), function(method) {
  gsea_df <- as.data.frame(gsea_results[[method]])
  hypoxia_row <- gsea_df[gsea_df$Description == "HALLMARK_HYPOXIA", ]
  
  if (nrow(hypoxia_row) == 1) {
    hypoxia_row <- hypoxia_row[, metrics]
    hypoxia_row$Method <- method
    return(hypoxia_row)
  } else {
    return(data.frame(matrix(NA, ncol = length(metrics) + 1,
                             dimnames = list(NULL, c(metrics, "Method")))))
  }
})

hypoxia_df <- bind_rows(hypoxia_summary)
rownames(hypoxia_df) <- c('RF', 'minProb', 'SoftHybrid', 'KNN', 'Gaussian')
```

```{r}
hypoxia_df <- hypoxia_df[,-c(1,3)]

protein_rf <- unlist(strsplit(gsea_results$RF@result$core_enrichment[1], split = "/"))
protein_minProb <- unlist(strsplit(gsea_results$minProb@result$core_enrichment[1], split = "/"))
protein_SoftHybrid <- unlist(strsplit(gsea_results$SoftHybrid@result$core_enrichment[1], split = "/"))
protein_KNN <- unlist(strsplit(gsea_results$KNN@result$core_enrichment[1], split = "/"))
protein_Gaussian <- unlist(strsplit(gsea_results$Gaussian@result$core_enrichment[1], split = "/"))

hypoxia_df$GeneRatio <- c(length(protein_rf), length(protein_minProb), length(protein_SoftHybrid), length(protein_KNN), length(protein_Gaussian))/200
```

```{r}
hypoxia_df$logFDR <- -log10(hypoxia_df$p.adjust)
hypoxia_df_all <- hypoxia_df %>%
  arrange(NES)
hypoxia_df_all$Method <- factor(hypoxia_df_all$Method, levels = hypoxia_df_all$Method)

pHypoxia_data <- 
ggplot(hypoxia_df_all, aes(x = NES, y = Method)) +
  geom_point(aes(size = GeneRatio, color = logFDR), shape = 16) +
  scale_color_gradient2(low = "#4575b4", mid = "white", high = "#d73027", midpoint = 2.5) +
  scale_size(range = c(3, 8), trans = "sqrt") +
  theme_bw() +
  labs(title = "GSEA Results for HALLMARK_HYPOXIA",
       x = "NES", y = "Imputation Method",
       size = "Gene Ratio", color = "-log10(adj.p-value)") +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pHypoxia_data
ggsave(
  filename = "./figure_benchmarking/Hypoxia_Results.tiff",  
  plot = pHypoxia_data,                      
  device = "tiff",                    
  dpi = 300,                         
  width = 10,                         
  height = 4,                       
  units = "in"                       
)
```
```{r}
hypoxia_df$logFDR <- -log10(hypoxia_df$p.adjust)
hypoxia_df_main <- hypoxia_df %>%
  filter(Method != 'KNN') %>%
  arrange(NES)
hypoxia_df_main$Method <- factor(hypoxia_df_main$Method, levels = hypoxia_df_main$Method)

pHypoxia_data <- 
ggplot(hypoxia_df_main, aes(x = NES, y = Method)) +
  geom_point(aes(size = GeneRatio, color = logFDR), shape = 16) +
  scale_color_gradient2(low = "#4575b4", mid = "white", high = "#d73027", midpoint = 2.5) +
  scale_size(range = c(3, 8), trans = "sqrt") +
  theme_bw() +
  labs(title = "GSEA Results for HALLMARK_HYPOXIA",
       x = "NES", y = "Imputation Method",
       size = "Gene Ratio", color = "-log10(adj.p-value)") +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15), 
    strip.text       = element_text(size = 13, face = "bold") 
  )
pHypoxia_data
ggsave(
  filename = "./figure_benchmarking/Hypoxia_Results_main.tiff",  
  plot = pHypoxia_data,                      
  device = "tiff",                    
  dpi = 300,                         
  width = 8,                         
  height = 4.5,                       
  units = "in"                       
)
```


```{r}
gsea_df <- data.frame(
  'Hallmark' = c(unlist(strsplit(gsea_results$RF@result$Description, split = '/', fixed = TRUE)),
  unlist(strsplit(gsea_results$minProb@result$Description, split = '/', fixed = TRUE)),
  unlist(strsplit(gsea_results$KNN@result$Description, split = '/', fixed = TRUE)),
  unlist(strsplit(gsea_results$Gaussian@result$Description, split = '/', fixed = TRUE))),
  'Method' = c(rep('RF', 7), rep('minProb', 8), rep('KNN', 7), rep('Gaussian', 8)),
  'pvalue' = c(gsea_results$RF@result$p.adjust, gsea_results$minProb@result$p.adjust, gsea_results$KNN@result$p.adjust, gsea_results$Gaussian@result$p.adjust),
  'GeneRatio' = c(pRF$data$GeneRatio, pMP$data$GeneRatio, pKNN$data$GeneRatio, pG$data$GeneRatio),
  'NES' = c(pRF$data$NES, pMP$data$NES, pKNN$data$NES, pG$data$NES),
  'Status' = 'activated'
)
```


```{r}
method_lv <- c("RF","minProb","KNN","Gaussian")
gsea_df$Method <- factor(gsea_df$Method, levels = intersect(method_lv, unique(gsea_df$Method)))
gsea_df <- gsea_df %>%
  mutate(logp = -log10(pvalue))
gsea_df$Status[gsea_df$NES < 0] <- "suppressed"

pathway_order <- gsea_df %>%
  group_by(Hallmark) %>%
  summarize(best_logp = max(logp, na.rm = TRUE), .groups = "drop") %>%
  arrange(best_logp) %>%
  pull(Hallmark)

gsea_df$Hallmark <- factor(gsea_df$Hallmark, levels = pathway_order)

gsea_df$Status <- factor(gsea_df$Status, levels = c("activated","suppressed"))

p_gsea_bubble <- ggplot(gsea_df, aes(x = Method, y = Hallmark,
                                     size = GeneRatio, color = logp)) +
  geom_point(alpha = 0.9) +
  scale_size(range = c(0.5, 8), name = "Gene ratio") +
  scale_color_viridis_c(direction = 1, name = expression(-log[10](pvalue))) +
  facet_grid(. ~ Status, scales = "free_x", space = "free_x") +  
  theme_bw() +
  labs(title = "GSEA bubble plot across imputation methods",
       x = "Imputation method", y = "Enriched pathways") +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(color = "black", size = 13, face = "bold"),
    axis.text.y      = element_text(color = "black", size = 13, face = "bold"), 
    axis.ticks.length= unit(2, "mm"),
    axis.title       = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position  = "right",
    legend.title     = element_text(size = 13, face = "bold"),
    legend.text      = element_text(size = 12),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    strip.text       = element_text(size = 13, face = "bold") 
  )
p_gsea_bubble
```
```{r}
ggsave(
  filename = "./figure_benchmarking/gsea_all_comboinone.tiff",  
  plot = p_gsea_bubble,                    
  device = "tiff",                    
  dpi = 300,                          
  width = 12.5,                        
  height = 5.5,                     
  units = "in"                        
)
```

# Hypxia related protein
## Missing rate
```{r}
hypoxia_genes <- hallmarks %>%
  filter(term == "HALLMARK_HYPOXIA") %>%
  pull(gene) %>%
  unique()

hypoxia_genes_in_data <- intersect(hypoxia_genes, rownames(df_filter))

length(hypoxia_genes_in_data)
hypoxia_genes_in_data[1:10]

df_hypoxia_raw <- df_filter[hypoxia_genes_in_data, , drop = FALSE]

hypoxia_samples  <- metadata$Sample[metadata$Condition == "Hypoxia"]
normoxia_samples <- metadata$Sample[metadata$Condition == "Normoxia"]

hypoxia_missing_df <- data.frame(
  Gene = rownames(df_hypoxia_raw),
  MissingRate_Overall  = rowMeans(is.na(df_hypoxia_raw)),
  MissingRate_Hypoxia  = rowMeans(is.na(df_hypoxia_raw[, hypoxia_samples, drop = FALSE])),
  MissingRate_Normoxia = rowMeans(is.na(df_hypoxia_raw[, normoxia_samples, drop = FALSE])),
  MeanAbundance_Observed = rowMeans(df_hypoxia_raw, na.rm = TRUE)
)

hypoxia_missing_df <- hypoxia_missing_df %>%
  arrange(desc(MissingRate_Overall))

head(hypoxia_missing_df)
```
```{r}
hypoxia_missing_df$Gene <- factor(
  hypoxia_missing_df$Gene,
  levels = hypoxia_missing_df$Gene[order(hypoxia_missing_df$MissingRate_Normoxia)]
)

hypoxia_missing_plot <- hypoxia_missing_df %>%
  dplyr::select(Gene, MissingRate_Normoxia, MissingRate_Hypoxia) %>%
  pivot_longer(
    cols = c(MissingRate_Normoxia, MissingRate_Hypoxia),
    names_to = "Condition",
    values_to = "MissingRate"
  ) %>%
  mutate(
    Condition = factor(
      Condition,
      levels = c("MissingRate_Normoxia", "MissingRate_Hypoxia"),
      labels = c("Normoxia", "Hypoxia")
    ),
    MissingRate_plot = ifelse(Condition == "Normoxia", -MissingRate, MissingRate)
  )

p_hypoxia_missing_bar <- ggplot(
  hypoxia_missing_plot,
  aes(x = Gene, y = MissingRate_plot, fill = Condition)
) +
  geom_col(width = 0.8) +
  coord_flip() +
  scale_y_continuous(
    labels = function(x) scales::percent(abs(x), accuracy = 1),
    breaks = seq(-1, 1, by = 0.25),
    limits = c(-1, 1)
  ) +
  scale_fill_manual(values = c("Normoxia" = "#0072B2", "Hypoxia" = "#D55E00")) +
  labs(
    title = "Missing rate of HALLMARK_HYPOXIA proteins in raw data",
    x = NULL,
    y = "Missing rate"
  ) +
  theme_bw() +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 11, face = "bold"),
    axis.text.y      = element_blank(),
    axis.ticks.y     = element_blank(),
    axis.title       = element_text(size = 14, face = "bold"),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.title     = element_blank(),
    legend.position  = "top"
  ) +
  geom_vline(xintercept = NULL) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black")

p_hypoxia_missing_bar
```

## Abundance Plot
```{r}
hypoxia_genes_in_data <- unique(c(protein_rf, protein_minProb, protein_SoftHybrid, protein_Gaussian))
```

```{r}
prepare_hypoxia_pathway_df <- function(expr_mat, metadata, method_name, gene_set) {
  
  expr_sub <- expr_mat[intersect(gene_set, rownames(expr_mat)), , drop = FALSE]
  
  df_long <- expr_sub %>%
    as.data.frame() %>%
    mutate(Gene = rownames(.)) %>%
    pivot_longer(
      cols = -Gene,
      names_to = "Sample",
      values_to = "Expression"
    ) %>%
    left_join(
      metadata %>% dplyr::select(Sample, Condition),
      by = "Sample"
    ) %>%
    mutate(Method = method_name)
  
  return(df_long)
}

plot_raw   <- prepare_hypoxia_pathway_df(df_norm,    metadata, "Raw",        hypoxia_genes_in_data)
plot_rf    <- prepare_hypoxia_pathway_df(imp_all,      metadata, "RF",         hypoxia_genes_in_data)
plot_mp    <- prepare_hypoxia_pathway_df(minProb_all,  metadata, "minProb",    hypoxia_genes_in_data)
plot_sh    <- prepare_hypoxia_pathway_df(df_imp_soft,  metadata, "SoftHybrid", hypoxia_genes_in_data)
plot_knn   <- prepare_hypoxia_pathway_df(knn_all,      metadata, "KNN",        hypoxia_genes_in_data)
plot_gs    <- prepare_hypoxia_pathway_df(GS_all,       metadata, "Gaussian",   hypoxia_genes_in_data)

hypoxia_plot_df <- bind_rows(
  plot_raw, plot_rf, plot_mp, plot_sh, plot_knn, plot_gs
)

hypoxia_plot_df$Condition <- factor(
  hypoxia_plot_df$Condition,
  levels = c("Normoxia", "Hypoxia")
)

hypoxia_plot_df$Method <- factor(
  hypoxia_plot_df$Method,
  levels = c("Raw", "RF", "minProb", "SoftHybrid", "KNN", "Gaussian")
)

hypoxia_mean_df <- hypoxia_plot_df %>%
  group_by(Method, Gene, Condition) %>%
  summarise(
    MeanExpression = mean(Expression, na.rm = TRUE),
    SD = sd(Expression, na.rm = TRUE),
    N = sum(!is.na(Expression)),
    SE = SD / sqrt(N),
    .groups = "drop"
  )

head(hypoxia_mean_df)
```

```{r}
p_hypoxia_diamond <- ggplot(
  hypoxia_mean_df,
  aes(x = Condition, y = MeanExpression, group = Gene)
) +
  geom_line(color = "grey60", alpha = 0.7, linewidth = 0.5) +
  geom_point(
    aes(fill = Condition),
    shape = 18,
    size = 2.6,
    color = "black",
    stroke = 0.3
  ) +
  facet_wrap(~Method, ncol = 3, scales = "free_y") +
  scale_fill_manual(values = c("Normoxia" = "#0072B2", "Hypoxia" = "#D55E00")) +
  labs(
    title = "HALLMARK_HYPOXIA proteins: Hypoxia vs Normoxia across imputation methods",
    x = NULL,
    y = "Mean log2 abundance"
  ) +
  theme_bw() +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    text             = element_text(size = 13),
    axis.text        = element_text(size = 12, face = "bold"),
    axis.title       = element_text(size = 14, face = "bold"),
    strip.text       = element_text(size = 13, face = "bold"),
    legend.position  = "none",
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

p_hypoxia_diamond
```
```{r}
hypoxia_gene_df <- hypoxia_mean_df %>%
  dplyr::select(Method, Gene, Condition, MeanExpression) %>%
  pivot_wider(
    names_from = Condition,
    values_from = MeanExpression
  ) %>%
  filter(!is.na(Normoxia) & !is.na(Hypoxia)) %>%
  mutate(
    delta = Hypoxia - Normoxia
  )

hypoxia_gene_long <- hypoxia_gene_df %>%
  group_by(Method) %>%
  arrange(delta, .by_group = TRUE) %>%
  mutate(Gene_order = factor(Gene, levels = unique(Gene))) %>%
  ungroup() %>%
  pivot_longer(
    cols = c(Normoxia, Hypoxia),
    names_to = "Condition",
    values_to = "MeanExpression"
  ) %>%
  mutate(
    Condition = factor(Condition, levels = c("Normoxia", "Hypoxia"))
  )

mean_line_df <- hypoxia_gene_long %>%
  group_by(Method, Condition) %>%
  summarise(
    MeanAll = mean(MeanExpression, na.rm = TRUE),
    .groups = "drop"
  )

p_hypoxia_dumbbell <- ggplot(hypoxia_gene_long,
                             aes(x = MeanExpression, y = Gene_order, group = Gene)) +
  geom_line(color = "grey70", linewidth = 0.6, alpha = 0.8) +
  geom_point(aes(fill = Condition),
             shape = 21, size = 2.8, color = "black", stroke = 0.3) +
  
  geom_vline(
    data = mean_line_df %>% filter(Condition == "Normoxia"),
    aes(xintercept = MeanAll),
    color = "#0072B2",
    linetype = "dashed",
    linewidth = 0.8,
    inherit.aes = FALSE
  ) +
  geom_vline(
    data = mean_line_df %>% filter(Condition == "Hypoxia"),
    aes(xintercept = MeanAll),
    color = "#D55E00",
    linetype = "dashed",
    linewidth = 0.8,
    inherit.aes = FALSE
  ) +
  
  facet_wrap(~Method, ncol = 3, scales = "free_y") +
  scale_fill_manual(values = c("Normoxia" = "#0072B2", "Hypoxia" = "#D55E00")) +
  labs(
    title = "Leading-edge HALLMARK_HYPOXIA proteins: paired abundance shifts",
    x = "Mean log2 abundance",
    y = NULL
  ) +
  theme_bw(base_size = 13) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    axis.text.x      = element_text(size = 11, face = "bold"),
    axis.text.y      = element_text(size = 9, face = "bold"),
    axis.title       = element_text(size = 14, face = "bold"),
    strip.text       = element_text(size = 13, face = "bold"),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    legend.position  = "top",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

p_hypoxia_dumbbell
```
```{r}
summary_df <- hypoxia_mean_df %>%
  group_by(Method, Condition) %>%
  summarise(
    Mean = mean(MeanExpression, na.rm = TRUE),
    SD   = sd(MeanExpression, na.rm = TRUE),
    N    = n(),
    SE   = SD / sqrt(N),
    .groups = "drop"
  )

summary_df$Method <- factor(
  summary_df$Method,
  levels = c("Raw", "RF", "minProb", "SoftHybrid", "KNN", "Gaussian")
)

summary_df$Condition <- factor(
  summary_df$Condition,
  levels = c("Normoxia", "Hypoxia")
)

summary_wide <- summary_df %>%
  dplyr::select(Method, Condition, Mean, SD) %>%
  pivot_wider(
    names_from = Condition,
    values_from = c(Mean, SD)
  )

library(ggplot2)

p_summary_dumbbell <- ggplot(summary_wide, aes(y = Method)) +
  geom_segment(
    aes(
      x = Mean_Normoxia,
      xend = Mean_Hypoxia,
      yend = Method
    ),
    color = "grey50",
    linewidth = 1.2
  ) +
  geom_point(
    aes(x = Mean_Normoxia),
    shape = 21,
    size = 3.5,
    fill = "#0072B2",
    color = "black",
    stroke = 0.4
  ) +
  geom_errorbarh(
    aes(
      xmin = Mean_Normoxia - SD_Normoxia,
      xmax = Mean_Normoxia + SD_Normoxia
    ),
    height = 0.2,
    color = "#0072B2",
    linewidth = 0.8
  ) +
  geom_point(
    aes(x = Mean_Hypoxia),
    shape = 21,
    size = 3.5,
    fill = "#D55E00",
    color = "black",
    stroke = 0.4
  ) +
  geom_errorbarh(
    aes(
      xmin = Mean_Hypoxia - SD_Hypoxia,
      xmax = Mean_Hypoxia + SD_Hypoxia
    ),
    height = 0.2,
    color = "#D55E00",
    linewidth = 0.8
  ) +
  
  labs(
    title = "Mean abundance shift of hypoxia proteins across methods",
    x = "Mean log2 abundance (± SD)",
    y = NULL
  ) +
  
  theme_bw(base_size = 13) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    axis.text        = element_text(size = 12, face = "bold"),
    axis.title       = element_text(size = 14, face = "bold"),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

p_summary_dumbbell
```
```{r}
summary_wide <- summary_df %>%
  dplyr::select(Method, Condition, Mean, SD) %>%
  pivot_wider(
    names_from = Condition,
    values_from = c(Mean, SD)
  ) %>%
  mutate(
    delta = Mean_Hypoxia - Mean_Normoxia,
    x_mid = (Mean_Hypoxia + Mean_Normoxia) / 2
  )

p_summary_dumbbell <- ggplot(summary_wide, aes(y = Method)) +
  geom_segment(
    aes(
      x = Mean_Normoxia,
      xend = Mean_Hypoxia,
      yend = Method
    ),
    color = "grey50",
    linewidth = 1.2
  ) +
  geom_point(
    aes(x = Mean_Normoxia),
    shape = 21,
    size = 3.5,
    fill = "#0072B2",
    color = "black",
    stroke = 0.4
  ) +
  geom_errorbarh(
    aes(
      xmin = Mean_Normoxia - SD_Normoxia,
      xmax = Mean_Normoxia + SD_Normoxia
    ),
    height = 0.2,
    color = "#0072B2",
    linewidth = 0.8
  ) +
  geom_point(
    aes(x = Mean_Hypoxia),
    shape = 21,
    size = 3.5,
    fill = "#D55E00",
    color = "black",
    stroke = 0.4
  ) +
  geom_errorbarh(
    aes(
      xmin = Mean_Hypoxia - SD_Hypoxia,
      xmax = Mean_Hypoxia + SD_Hypoxia
    ),
    height = 0.2,
    color = "#D55E00",
    linewidth = 0.8
  ) +
  geom_text(
    aes(
      x = x_mid,
      label = sprintf("\u0394 = %.2f", delta)
    ),
    nudge_y = 0.28,
    size = 4,
    fontface = "bold"
  ) +
  labs(
    title = "Mean abundance shift of hypoxia proteins across methods",
    x = "Mean log2 abundance (\u00b1 SD)",
    y = NULL
  ) +
  theme_bw(base_size = 13) +
  theme(
    panel.border     = element_rect(colour = "black", fill = NA, linewidth = 0.6),
    axis.text        = element_text(size = 12, face = "bold"),
    axis.title       = element_text(size = 14, face = "bold"),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 15),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

p_summary_dumbbell
```


# Export
```{r}
ggsave(
  filename = "./figure_benchmarking/HALLMARK_HYPOXIA_missing_rate_raw_bar.tiff",
  plot = p_hypoxia_missing_bar,
  device = "tiff",
  dpi = 300,
  width = 8,
  height = 4.5,
  units = "in"
)

ggsave(
  filename = "./figure_benchmarking/HALLMARK_HYPOXIA_p_hypoxia_dumbbell.tiff",
  plot = p_hypoxia_dumbbell,
  device = "tiff",
  dpi = 300,
  width = 8,
  height = 10,
  units = "in"
)

ggsave(
  filename = "./figure_benchmarking/HALLMARK_HYPOXIA_p_summary_dumbbell.tiff",
  plot = p_summary_dumbbell,
  device = "tiff",
  dpi = 300,
  width = 8,
  height = 4.5,
  units = "in"
)
```



```{r}
sessionInfo()

```

























































